---
title: "Tutorial 8: Bill Success Factors"
subtitle: "assemblykor - Korean National Assembly Data for Political Science Education"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(assemblykor)
library(dplyr)
library(ggplot2)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
data(bills)
data(legislators)
set_ko_font()
tutorial_options(exercise.eval = FALSE)
```

```{r theme, echo=FALSE}
htmltools::includeCSS(
  system.file("tutorials/css/assemblykor.css", package = "assemblykor")
)
```

## Learning Objectives

**Research Question**: What determines whether a bill passes in the Korean National Assembly?

Tens of thousands of bills are proposed each term, but only a fraction actually pass. What factors determine success? Proposer seniority? Party? Committee?

Since the outcome is binary (pass/fail), we use **logistic regression** instead of OLS.

You will learn:

- Exploring bill processing outcomes
- Why logistic regression is needed for binary outcomes
- Running logistic regression with `glm()`
- Interpreting odds ratios
- Visualizing predicted probabilities

## 1. Bill Processing Outcomes

```{r result-overview, exercise=TRUE}
bills %>%
  count(result, sort = TRUE) %>%
  mutate(pct = round(n / sum(n) * 100, 1)) %>%
  head(10)
```

Most bills die from "term expiration" (the assembly term ends before they are processed). This is a distinctive feature of the Korean National Assembly.

### Creating a Binary Variable

```{r binary, exercise=TRUE}
bill_analysis <- bills %>%
  mutate(
    passed = as.integer(result %in% c("\uc6d0\uc548\uac00\uacb0", "\uc218\uc815\uac00\uacb0", "\ub300\uc548\ubc18\uc601\ud3d0\uae30")),
    year = as.integer(format(propose_date, "%Y"))
  )

cat("Total bills:", nrow(bill_analysis), "\n")
cat("Passed:", sum(bill_analysis$passed), "\n")
cat("Pass rate:", round(mean(bill_analysis$passed) * 100, 1), "%\n")
```

```{r quiz-binary}
quiz(
  question("A researcher codes 'alternative-reflected disposal' as failure (0). Another codes it as success (1). Who is more defensible?",
    answer("Failure, because the original bill was officially discarded"),
    answer("Success, because the bill's content was incorporated into a passed alternative", correct = TRUE),
    answer("Neither, because it depends on the research question"),
    answer("Both are equally valid regardless of context"),
    correct = "Correct! The bill's substance succeeded. Coding it as failure would undercount policy impact, since the content lives on in the alternative bill."
  )
)
```

### Pass Rate by Assembly

```{r pass-rate-assembly, exercise=TRUE}
bill_analysis <- bills %>%
  mutate(passed = as.integer(result %in% c("\uc6d0\uc548\uac00\uacb0", "\uc218\uc815\uac00\uacb0", "\ub300\uc548\ubc18\uc601\ud3d0\uae30")))

bill_analysis %>%
  group_by(assembly) %>%
  summarise(
    n = n(),
    n_passed = sum(passed),
    pass_rate = round(mean(passed) * 100, 1)
  )
```

## 2. Proposer Characteristics and Bill Success

```{r merge-legislator, exercise=TRUE}
bill_analysis <- bills %>%
  mutate(passed = as.integer(result %in% c("\uc6d0\uc548\uac00\uacb0", "\uc218\uc815\uac00\uacb0", "\ub300\uc548\ubc18\uc601\ud3d0\uae30")))

bill_leg <- bill_analysis %>%
  inner_join(
    legislators %>% select(member_id, assembly, party, gender,
                           seniority, district_type),
    by = c("proposer_id" = "member_id", "assembly")
  )

cat("Analysis sample:", nrow(bill_leg), "bills (with proposer info)\n")
```

### Pass Rate by Seniority

```{r pass-by-seniority, exercise=TRUE}
bill_analysis <- bills %>%
  mutate(passed = as.integer(result %in% c("\uc6d0\uc548\uac00\uacb0", "\uc218\uc815\uac00\uacb0", "\ub300\uc548\ubc18\uc601\ud3d0\uae30")))

bill_leg <- bill_analysis %>%
  inner_join(
    legislators %>% select(member_id, assembly, party, gender,
                           seniority, district_type),
    by = c("proposer_id" = "member_id", "assembly")
  )

pass_senior <- bill_leg %>%
  group_by(seniority) %>%
  summarise(n = n(), pass_rate = mean(passed)) %>%
  filter(n >= 50)

ggplot(pass_senior, aes(x = seniority, y = pass_rate)) +
  geom_col(fill = "steelblue", width = 0.6) +
  geom_text(aes(label = paste0(round(pass_rate * 100, 1), "%")),
            vjust = -0.5, size = 3.5) +
  scale_y_continuous(labels = scales::percent,
                     expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Seniority (terms elected)", y = "Bill Pass Rate",
       title = "Bill Pass Rate by Proposer Seniority") +
  theme_minimal()
```

### Pass Rate by Committee

```{r pass-by-committee, exercise=TRUE}
bill_analysis <- bills %>%
  mutate(passed = as.integer(result %in% c("\uc6d0\uc548\uac00\uacb0", "\uc218\uc815\uac00\uacb0", "\ub300\uc548\ubc18\uc601\ud3d0\uae30")))

bill_leg <- bill_analysis %>%
  inner_join(
    legislators %>% select(member_id, assembly, party, gender,
                           seniority, district_type),
    by = c("proposer_id" = "member_id", "assembly")
  )

pass_committee <- bill_leg %>%
  group_by(committee) %>%
  summarise(n = n(), pass_rate = mean(passed)) %>%
  filter(n >= 100) %>%
  arrange(desc(pass_rate))

ggplot(pass_committee %>% head(15),
       aes(x = reorder(committee, pass_rate), y = pass_rate)) +
  geom_col(fill = "coral") +
  geom_text(aes(label = paste0(round(pass_rate * 100, 1), "%")),
            hjust = -0.1, size = 3) +
  coord_flip() +
  scale_y_continuous(labels = scales::percent,
                     expand = expansion(mult = c(0, 0.15))) +
  labs(x = NULL, y = "Pass Rate",
       title = "Bill Pass Rate by Committee (Top 15)") +
  theme_minimal()
```

## 3. Why Logistic Regression?

When the outcome is 0 or 1, OLS has problems:

```{r ols-problem, exercise=TRUE}
bill_analysis <- bills %>%
  mutate(passed = as.integer(result %in% c("\uc6d0\uc548\uac00\uacb0", "\uc218\uc815\uac00\uacb0", "\ub300\uc548\ubc18\uc601\ud3d0\uae30")))

bill_leg <- bill_analysis %>%
  inner_join(
    legislators %>% select(member_id, assembly, party, gender,
                           seniority, district_type),
    by = c("proposer_id" = "member_id", "assembly")
  )

ggplot(bill_leg, aes(x = seniority, y = passed)) +
  geom_jitter(height = 0.05, alpha = 0.05, color = "gray50") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"),
              color = "blue", se = FALSE) +
  labs(
    x = "Seniority", y = "Passed (0 = No, 1 = Yes)",
    title = "OLS (red) vs Logistic (blue) Comparison"
  ) +
  theme_minimal()
```

- **Red line (OLS)**: Straight line that can exceed 0-1 range
- **Blue line (Logistic)**: S-curve always between 0 and 1

**Logistic regression** models log-odds instead of probability directly:

$$\log\left(\frac{p}{1-p}\right) = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots$$

```{r quiz-logistic}
quiz(
  question("Your OLS model predicts a bill's passage probability as -0.12. What does this reveal?",
    answer("The bill is very unlikely to pass"),
    answer("The linear model produced an impossible probability, which is why logistic regression is preferred", correct = TRUE),
    answer("There is a data entry error"),
    answer("The coefficient should be exponentiated"),
    correct = "Correct! Probabilities must be between 0 and 1. A negative prediction shows the fundamental limitation of OLS for binary outcomes."
  )
)
```

## 4. Running Logistic Regression

In R, use `glm()` with `family = binomial`:

```{r logit-simple, exercise=TRUE}
bill_analysis <- bills %>%
  mutate(passed = as.integer(result %in% c("\uc6d0\uc548\uac00\uacb0", "\uc218\uc815\uac00\uacb0", "\ub300\uc548\ubc18\uc601\ud3d0\uae30")))

bill_leg <- bill_analysis %>%
  inner_join(
    legislators %>% select(member_id, assembly, party, gender,
                           seniority, district_type),
    by = c("proposer_id" = "member_id", "assembly")
  )

m1 <- glm(passed ~ seniority, data = bill_leg, family = binomial)
summary(m1)
```

### Odds Ratios

Logistic regression coefficients are in **log-odds** units. Exponentiate to get **odds ratios**:

```{r odds-ratio, exercise=TRUE}
bill_analysis <- bills %>%
  mutate(passed = as.integer(result %in% c("\uc6d0\uc548\uac00\uacb0", "\uc218\uc815\uac00\uacb0", "\ub300\uc548\ubc18\uc601\ud3d0\uae30")))

bill_leg <- bill_analysis %>%
  inner_join(
    legislators %>% select(member_id, assembly, party, gender,
                           seniority, district_type),
    by = c("proposer_id" = "member_id", "assembly")
  )

m1 <- glm(passed ~ seniority, data = bill_leg, family = binomial)
exp(coef(m1))
```

> **Interpreting odds ratios**: When seniority increases by 1, the odds of passing multiply by this factor.
> - OR = 1: no effect
> - OR > 1: higher odds of passing
> - OR < 1: lower odds of passing

## 5. Multiple Logistic Regression

Control for multiple factors simultaneously:

```{r logit-multiple, exercise=TRUE}
bill_analysis <- bills %>%
  mutate(passed = as.integer(result %in% c("\uc6d0\uc548\uac00\uacb0", "\uc218\uc815\uac00\uacb0", "\ub300\uc548\ubc18\uc601\ud3d0\uae30")))

bill_leg <- bill_analysis %>%
  inner_join(
    legislators %>% select(member_id, assembly, party, gender,
                           seniority, district_type),
    by = c("proposer_id" = "member_id", "assembly")
  )

bill_leg_model <- bill_leg %>%
  filter(party %in% c("\uad6d\ubbfc\uc758\ud798", "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9", "\uc870\uad6d\ud601\uc2e0\ub2f9")) %>%
  mutate(party = relevel(factor(party), ref = "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9"))

m2 <- glm(passed ~ seniority + gender + district_type + party,
           data = bill_leg_model, family = binomial)
summary(m2)
```

### Odds Ratio Visualization

```{r or-plot, exercise=TRUE}
bill_analysis <- bills %>%
  mutate(passed = as.integer(result %in% c("\uc6d0\uc548\uac00\uacb0", "\uc218\uc815\uac00\uacb0", "\ub300\uc548\ubc18\uc601\ud3d0\uae30")))

bill_leg <- bill_analysis %>%
  inner_join(
    legislators %>% select(member_id, assembly, party, gender,
                           seniority, district_type),
    by = c("proposer_id" = "member_id", "assembly")
  )

bill_leg_model <- bill_leg %>%
  filter(party %in% c("\uad6d\ubbfc\uc758\ud798", "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9", "\uc870\uad6d\ud601\uc2e0\ub2f9")) %>%
  mutate(party = relevel(factor(party), ref = "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9"))

m2 <- glm(passed ~ seniority + gender + district_type + party,
           data = bill_leg_model, family = binomial)

or_df <- broom::tidy(m2, conf.int = TRUE, exponentiate = TRUE) %>%
  filter(term != "(Intercept)")

ggplot(or_df, aes(x = estimate, y = reorder(term, estimate))) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high),
                  color = "steelblue", size = 0.8) +
  labs(
    x = "Odds Ratio (95% CI)",
    y = NULL,
    title = "Determinants of Bill Passage (Logistic Regression)"
  ) +
  theme_minimal(base_size = 13)
```

- Point right of 1: **increases** odds of passage
- Point left of 1: **decreases** odds of passage
- CI crossing 1: not statistically significant

## 6. Predicted Probabilities

Converting logistic results to **predicted probabilities** makes them more intuitive.

```{r predicted-prob, exercise=TRUE}
bill_analysis <- bills %>%
  mutate(passed = as.integer(result %in% c("\uc6d0\uc548\uac00\uacb0", "\uc218\uc815\uac00\uacb0", "\ub300\uc548\ubc18\uc601\ud3d0\uae30")))

bill_leg <- bill_analysis %>%
  inner_join(
    legislators %>% select(member_id, assembly, party, gender,
                           seniority, district_type),
    by = c("proposer_id" = "member_id", "assembly")
  )

bill_leg_model <- bill_leg %>%
  filter(party %in% c("\uad6d\ubbfc\uc758\ud798", "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9", "\uc870\uad6d\ud601\uc2e0\ub2f9")) %>%
  mutate(party = relevel(factor(party), ref = "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9"))

m2 <- glm(passed ~ seniority + gender + district_type + party,
           data = bill_leg_model, family = binomial)

# Predicted probabilities by seniority and party
new_party <- expand.grid(
  seniority = 1:5,
  gender = "M",
  district_type = "constituency",
  party = factor(c("\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9", "\uad6d\ubbfc\uc758\ud798", "\uc870\uad6d\ud601\uc2e0\ub2f9"),
                 levels = levels(bill_leg_model$party))
)

new_party$pred_prob <- predict(m2, newdata = new_party, type = "response")

ggplot(new_party, aes(x = seniority, y = pred_prob, color = party)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = c(
    "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9" = "#004EA2",
    "\uad6d\ubbfc\uc758\ud798" = "#E61E2B",
    "\uc870\uad6d\ud601\uc2e0\ub2f9" = "#00B0F0"
  )) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Seniority", y = "Predicted Pass Probability",
       color = "Party",
       title = "Predicted Bill Pass Probability by Party and Seniority") +
  theme_minimal()
```

## 7. Model Comparison

```{r model-compare, exercise=TRUE}
bill_analysis <- bills %>%
  mutate(passed = as.integer(result %in% c("\uc6d0\uc548\uac00\uacb0", "\uc218\uc815\uac00\uacb0", "\ub300\uc548\ubc18\uc601\ud3d0\uae30")))

bill_leg <- bill_analysis %>%
  inner_join(
    legislators %>% select(member_id, assembly, party, gender,
                           seniority, district_type),
    by = c("proposer_id" = "member_id", "assembly")
  )

bill_leg_model <- bill_leg %>%
  filter(party %in% c("\uad6d\ubbfc\uc758\ud798", "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9", "\uc870\uad6d\ud601\uc2e0\ub2f9")) %>%
  mutate(party = relevel(factor(party), ref = "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9"))

m1 <- glm(passed ~ seniority, data = bill_leg_model, family = binomial)
m2 <- glm(passed ~ seniority + gender + district_type + party,
           data = bill_leg_model, family = binomial)
m3 <- glm(passed ~ seniority + gender + district_type + party + assembly,
           data = bill_leg_model, family = binomial)

data.frame(
  Model = c("M1: Seniority only", "M2: + Individual traits", "M3: + Assembly"),
  AIC = c(round(AIC(m1)), round(AIC(m2)), round(AIC(m3))),
  Pseudo_R2 = c(
    round(1 - m1$deviance / m1$null.deviance, 4),
    round(1 - m2$deviance / m2$null.deviance, 4),
    round(1 - m3$deviance / m3$null.deviance, 4)
  )
)
```

> **AIC**: Lower is better. Balances model fit and complexity.

> **Pseudo R-squared**: Analogous to OLS R-squared. Higher means better explanatory power, but values tend to be lower than OLS R-squared.

## Summary

| Concept | Description |
|---------|-------------|
| Binary DV | Outcome is 0/1 (pass/fail) |
| Logistic regression | `glm(..., family = binomial)` |
| Odds ratio (OR) | `exp(coefficient)`, >1 means higher odds |
| Predicted probability | `predict(..., type = "response")` |
| AIC | Model comparison metric (lower = better) |

## Exercises

### Exercise 1: Gender Effect

Run a logistic regression with only `gender` as predictor. What is the odds ratio for male vs female?

```{r ex1, exercise=TRUE}
# Your code here

```

```{r ex1-solution}
bill_analysis <- bills %>%
  mutate(passed = as.integer(result %in% c("\uc6d0\uc548\uac00\uacb0", "\uc218\uc815\uac00\uacb0", "\ub300\uc548\ubc18\uc601\ud3d0\uae30")))

bill_leg <- bill_analysis %>%
  inner_join(
    legislators %>% select(member_id, assembly, party, gender,
                           seniority, district_type),
    by = c("proposer_id" = "member_id", "assembly")
  )

m_gender <- glm(passed ~ gender, data = bill_leg, family = binomial)
exp(coef(m_gender))
```

### Exercise 2: Committee Effects

Add `committee` to the model. Which 3 committees have the highest odds of bill passage? (Use only the top 10 committees by bill count.)

```{r ex2, exercise=TRUE}
# Your code here

```

```{r ex2-hint}
# Filter to top 10 committees first
# Then run glm with committee as a factor variable
# Use broom::tidy(model, exponentiate = TRUE) to get odds ratios
```

### Exercise 3: Co-sponsorship

Does the number of co-sponsors affect bill passage? Use `get_proposers()` to count co-sponsors per bill and test with logistic regression.

```{r ex3, exercise=TRUE}
# This requires downloading proposer data first
# proposers <- get_proposers()
# Then count co-sponsors per bill and merge with bill data

```

```{r ex3-hint}
# proposers <- get_proposers()
# n_cosponsors <- proposers %>%
#   filter(!is_lead) %>%
#   count(bill_id, name = "n_cosponsors")
# bill_cosponsor <- bill_analysis %>%
#   inner_join(n_cosponsors, by = "bill_id")
# m <- glm(passed ~ n_cosponsors, data = bill_cosponsor, family = binomial)
```
