---
title: "Data Codebook"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Codebook}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(assemblykor)

# Helper: summarise a single variable
var_summary <- function(x, varname) {
  n_total <- length(x)
  n_miss  <- sum(is.na(x))
  pct_miss <- sprintf("%.1f%%", 100 * n_miss / n_total)

  if (is.logical(x)) {
    type_str <- "logical"
    vals <- paste0("TRUE: ", sum(x, na.rm = TRUE),
                   ", FALSE: ", sum(!x, na.rm = TRUE))
  } else if (is.numeric(x)) {
    type_str <- "numeric"
    q <- quantile(x, c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
    vals <- sprintf("min=%.0f, Q1=%.0f, median=%.0f, Q3=%.0f, max=%.0f",
                    q[1], q[2], q[3], q[4], q[5])
  } else if (inherits(x, "Date")) {
    type_str <- "Date"
    vals <- paste(range(x, na.rm = TRUE), collapse = " to ")
  } else {
    type_str <- "character"
    u <- length(unique(x[!is.na(x)]))
    top <- names(sort(table(x), decreasing = TRUE))[1:min(3, u)]
    vals <- paste0(u, " unique; top: ", paste(top, collapse = ", "))
  }

  data.frame(
    Variable = varname,
    Type = type_str,
    Missing = pct_miss,
    Distribution = vals,
    stringsAsFactors = FALSE
  )
}

# Helper: build a summary table for a dataset
codebook_table <- function(df) {
  rows <- lapply(names(df), function(v) var_summary(df[[v]], v))
  do.call(rbind, rows)
}
```

This codebook documents all seven built-in datasets in `assemblykor`.
For each dataset, we list every variable with its type, missing rate,
and value distribution. All datasets can be joined via `member_id`
and/or `assembly`.

---

## legislators

**947 rows, 15 variables.** MP metadata for the 20th-22nd Korean
National Assembly.

- **Unit of observation**: legislator-assembly
- **Key**: `member_id` + `assembly` (unique)
- **Source**: Open National Assembly API

```{r legislators-codebook, echo = FALSE}
data(legislators)
knitr::kable(codebook_table(legislators), row.names = FALSE)
```

---

## bills

**60,925 rows, 9 variables.** Legislative bill metadata (20th-22nd
assembly).

- **Unit of observation**: bill
- **Key**: `bill_id` (unique)
- **Join**: `proposer_id` links to `legislators$member_id`
- **Source**: Open National Assembly API

```{r bills-codebook, echo = FALSE}
data(bills)
knitr::kable(codebook_table(bills), row.names = FALSE)
```

---

## wealth

**2,928 rows, 14 variables.** Legislator asset declaration panel
(2015-2025, 13 disclosure periods).

- **Unit of observation**: legislator-year
- **Key**: `member_id` + `year` (unique)
- **Units**: all monetary values in thousands of KRW (1 unit = 1,000 won)
- **Source**: OpenWatch (CC BY-SA 4.0)

```{r wealth-codebook, echo = FALSE}
data(wealth)
knitr::kable(codebook_table(wealth), row.names = FALSE)
```

---

## seminars

**5,962 rows, 18 variables.** Legislator-year policy seminar activity
(17th-22nd assembly, 2000-2025).

- **Unit of observation**: legislator-year
- **Key**: `member_id` + `year` (note: ~5% of `member_id` are `NA`)
- **Source**: National Assembly Seminar Database

```{r seminars-codebook, echo = FALSE}
data(seminars)
knitr::kable(codebook_table(seminars), row.names = FALSE)
```

---

## speeches

**15,843 rows, 9 variables.** Committee speech records from the
Science and ICT Committee (22nd assembly, 2024).

- **Unit of observation**: speech turn
- **Key**: `date` + `speech_order` (unique within a meeting)
- **Source**: National Assembly committee minutes

```{r speeches-codebook, echo = FALSE}
data(speeches)
knitr::kable(codebook_table(speeches), row.names = FALSE)
```

---

## votes

**7,997 rows, 13 variables.** Plenary vote tallies (20th-22nd
assembly).

- **Unit of observation**: bill vote
- **Key**: `bill_id` (unique)
- **Join**: `bill_id` links to `bills$bill_id` (~40% match rate; `votes`
  includes committee alternatives and budget bills not in `bills`)
- **Source**: Open National Assembly API

```{r votes-codebook, echo = FALSE}
data(votes)
knitr::kable(codebook_table(votes), row.names = FALSE)
```

---

## roll_calls

**368,210 rows, 8 variables.** Member-level roll call votes (22nd
assembly, 1,233 bills).

- **Unit of observation**: legislator-bill vote
- **Key**: `member_id` + `bill_id` (unique)
- **Join**: `member_id` links to `legislators$member_id`;
  `bill_id` links to `votes$bill_id`
- **Source**: Open National Assembly API

```{r roll-calls-codebook, echo = FALSE}
data(roll_calls)
knitr::kable(codebook_table(roll_calls), row.names = FALSE)
```

---

## Dataset relationship diagram

```
                    legislators
                   (member_id + assembly)
                   /       |        \
                  /        |         \
               wealth   seminars   bills
            (member_id) (member_id) (proposer_id)
                                      |
                                    votes
                                  (bill_id)
                                      |
                                  roll_calls
                               (bill_id + member_id)
                                      |
                                  legislators
                                  (member_id)

    speeches --- legislators (member_id, 22nd assembly only)
```

All datasets share `member_id` as the primary join key. Use
`assembly` as a secondary key when joining datasets that span
multiple assembly terms.
