---
title: "5-Minute Quickstart"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{5-Minute Quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4
)
```

## Install and load

```r
# install.packages("remotes")
remotes::install_github("kyusik-yang/assemblykor")
```

```{r load}
library(assemblykor)
```

## Seven datasets, one line each

```{r overview}
# All datasets load lazily - just call their name
head(legislators, 3)
head(bills, 3)
head(wealth, 3)
```

## Quick analysis: gender and legislative productivity

```{r gender, message = FALSE}
library(dplyr)

legislators %>%
  filter(assembly == 22) %>%
  group_by(gender) %>%
  summarise(
    n = n(),
    median_bills_led = median(n_bills_lead, na.rm = TRUE)
  )
```

## Quick analysis: bill survival rates

```{r bills-survival}
bills %>%
  count(result, sort = TRUE) %>%
  head(5)
```

Only about 5% of bills pass in their original form.

## Quick analysis: wealth inequality

```{r wealth-ineq}
# Net worth in billion KRW
hist(wealth$net_worth / 1e6, breaks = 50, col = "steelblue",
     main = "Legislator Net Worth", xlab = "Billion KRW")
```

## Quick analysis: voting consensus

```{r votes-consensus}
votes$yes_rate <- votes$yes / votes$voted
hist(votes$yes_rate, breaks = 40, col = "coral",
     main = "Distribution of Yes-Vote Share", xlab = "Proportion yes")
```

Most bills pass near-unanimously; a handful are fiercely contested.

## Joining datasets

All datasets share `member_id` for easy merging:

```{r join, message = FALSE}
# Do wealthier legislators propose more bills?
leg_wealth <- legislators %>%
  filter(assembly == 22) %>%
  inner_join(wealth %>% filter(year == 2024),
             by = "member_id")

cor(leg_wealth$n_bills_lead, leg_wealth$net_worth,
    use = "complete.obs")
```

## Tutorials

Nine Korean-language tutorials cover the full methods sequence:

```{r tutorials, eval = FALSE}
list_tutorials()           # See all 9
run_tutorial(1)            # Interactive in browser
open_tutorial(1)           # Copy Rmd to your directory
```

## Next steps

- `vignette("introduction")` for deeper examples
- `vignette("codebook")` for the full data dictionary
- `?legislators`, `?bills`, etc. for variable-level documentation
