---
title: "Tutorial 9: Speech Pattern Analysis"
subtitle: "assemblykor - Korean National Assembly Data for Political Science Education"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(assemblykor)
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
data(speeches)
data(legislators)
set_ko_font()
tutorial_options(exercise.eval = FALSE)
```

```{r theme, echo=FALSE}
htmltools::includeCSS(
  system.file("tutorials/css/assemblykor.css", package = "assemblykor")
)
```

## Learning Objectives

**Research Question**: Who speaks, how much, and about what in National Assembly committees?

Standing committees are the core arena for bill deliberation. Legislators scrutinize the government through questions, and bureaucrats explain policies through answers. Speech data contains rich information about participation patterns, agenda-setting, and question-answer dynamics.

This tutorial focuses on **descriptive statistics and visualization**. You can discover meaningful patterns without complex models.

You will learn:

- Speech data structure and speaker roles
- Participation measurement: who speaks most and longest?
- Question-answer turn structure
- Agenda-specific speech analysis
- Speech diversity measurement

## 1. Data Structure

The `speeches` data contains all speeches from the 22nd Assembly's Science, ICT, Broadcasting and Communications Committee.

```{r overview, exercise=TRUE}
cat("Total speeches:", nrow(speeches), "\n")
cat("Meeting dates:", n_distinct(speeches$date), "\n")
cat("Speakers:", n_distinct(speeches$speaker_name), "\n")
```

### Speaker Roles

Various roles participate in committee proceedings:

```{r roles, exercise=TRUE}
speeches %>%
  count(role, sort = TRUE) %>%
  mutate(pct = round(n / sum(n) * 100, 1))
```

```{r role-plot, exercise=TRUE}
role_summary <- speeches %>%
  count(role, sort = TRUE)

ggplot(role_summary, aes(x = reorder(role, n), y = n)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(x = NULL, y = "Speech Count",
       title = "Speech Count by Speaker Role (22nd, Sci & ICT Committee)") +
  theme_minimal()
```

```{r quiz-roles, echo=FALSE}
quiz(
  question("The data shows legislators give far more speeches than ministers. Does this mean legislators contribute more substantive content?",
    answer("Yes, higher speech count directly indicates more substance"),
    answer("Not necessarily, because many of those are brief questions", correct = TRUE),
    answer("Yes, because elected officials have greater policy authority"),
    answer("No, because government officials always provide more depth"),
    correct = "Correct! Speech count alone doesn't capture substance. Legislators often give many short turns (questions, procedural remarks), while ministers give fewer but longer policy responses."
  )
)
```

## 2. Participation Analysis

### Speech Frequency

```{r freq-top, exercise=TRUE}
leg_speeches <- speeches %>%
  filter(role == "legislator")

top_speakers <- leg_speeches %>%
  count(speaker_name, sort = TRUE) %>%
  head(15)

ggplot(top_speakers, aes(x = reorder(speaker_name, n), y = n)) +
  geom_col(fill = "coral") +
  coord_flip() +
  labs(x = NULL, y = "Speech Count",
       title = "Most Frequent Speakers (22nd, Sci & ICT, Top 15)") +
  theme_minimal()
```

### Speech Length

Speech length matters as much as frequency. Some legislators ask short questions frequently, others deliver long statements.

```{r length-analysis, exercise=TRUE}
leg_speeches <- speeches %>%
  filter(role == "legislator") %>%
  mutate(speech_length = nchar(speech))

speaker_stats <- leg_speeches %>%
  group_by(speaker_name) %>%
  summarise(
    n_speeches = n(),
    total_chars = sum(speech_length),
    avg_length = round(mean(speech_length)),
    median_length = median(speech_length)
  ) %>%
  filter(n_speeches >= 10) %>%
  arrange(desc(total_chars))

head(speaker_stats, 10)
```

### Frequency vs Length

```{r freq-vs-length, exercise=TRUE}
leg_speeches <- speeches %>%
  filter(role == "legislator") %>%
  mutate(speech_length = nchar(speech))

speaker_stats <- leg_speeches %>%
  group_by(speaker_name) %>%
  summarise(
    n_speeches = n(),
    avg_length = round(mean(speech_length))
  ) %>%
  filter(n_speeches >= 10)

ggplot(speaker_stats, aes(x = n_speeches, y = avg_length)) +
  geom_point(alpha = 0.6, color = "steelblue", size = 2) +
  labs(
    x = "Number of Speeches", y = "Average Speech Length (characters)",
    title = "Speech Frequency vs Average Length"
  ) +
  theme_minimal()
```

Do frequent speakers also speak longer, or do they tend to ask short questions?

## 3. Role-based Speech Patterns

The question-answer structure of committees creates distinct patterns by role.

```{r role-length, exercise=TRUE}
speeches %>%
  mutate(
    speech_length = nchar(speech),
    role_group = case_match(role,
      "legislator" ~ "Legislator",
      "chair" ~ "Chair",
      c("minister", "vice_minister", "senior_bureaucrat") ~ "Government",
      "committee_staff" ~ "Staff",
      .default = "Other"
    )
  ) %>%
  group_by(role_group) %>%
  summarise(
    n = n(),
    avg_length = round(mean(speech_length)),
    median_length = round(median(speech_length)),
    total_chars = sum(speech_length)
  ) %>%
  arrange(desc(total_chars))
```

### Length Distribution by Role

```{r role-density, exercise=TRUE}
speeches %>%
  mutate(
    speech_length = nchar(speech),
    role_group = case_match(role,
      "legislator" ~ "Legislator",
      c("minister", "vice_minister", "senior_bureaucrat") ~ "Government",
      .default = NA_character_
    )
  ) %>%
  filter(!is.na(role_group)) %>%
  ggplot(aes(x = speech_length, fill = role_group)) +
  geom_density(alpha = 0.5) +
  scale_x_log10(labels = scales::comma) +
  scale_fill_manual(values = c("Legislator" = "steelblue", "Government" = "coral")) +
  labs(
    x = "Speech Length (characters, log scale)",
    y = "Density", fill = NULL,
    title = "Legislator vs Government Speech Length Distribution"
  ) +
  theme_minimal()
```

```{r quiz-role, echo=FALSE}
quiz(
  question("The density plot shows government officials' speeches are much longer than legislators'. If you only used mean speech length to measure 'engagement', what would be misleading?",
    answer("Nothing, because speech length directly measures effort"),
    answer("Officials would look more engaged, ignoring legislators' many brief questions", correct = TRUE),
    answer("It would systematically favor junior legislators over seniors"),
    answer("The density plot cannot be used for any kind of comparison"),
    correct = "Correct! The Q&A structure means legislators ask brief questions and officials give long answers. Using only length would mistake institutional roles for engagement levels."
  )
)
```

## 4. Turn Structure

Use `speech_order` to track the sequence of speeches within a meeting.

```{r turn-structure, exercise=TRUE}
sample_date <- speeches %>%
  count(date, sort = TRUE) %>%
  head(1) %>%
  pull(date)

turns <- speeches %>%
  filter(date == sample_date) %>%
  arrange(speech_order) %>%
  mutate(
    role_group = case_match(role,
      "legislator" ~ "Legislator",
      "chair" ~ "Chair",
      c("minister", "vice_minister", "senior_bureaucrat",
        "agency_head") ~ "Government",
      .default = "Other"
    ),
    speech_length = nchar(speech)
  ) %>%
  head(40)

ggplot(turns, aes(x = speech_order, y = speech_length, fill = role_group)) +
  geom_col(width = 0.8) +
  scale_fill_manual(values = c(
    "Legislator" = "steelblue", "Government" = "coral",
    "Chair" = "gray70", "Other" = "gray90"
  )) +
  labs(
    x = "Speech Order", y = "Speech Length (characters)",
    fill = "Role",
    title = paste("Meeting Turn Structure (", sample_date, ")")
  ) +
  theme_minimal()
```

Notice the alternating pattern of legislator (blue) and government (red) speeches, reflecting the Q&A format.

## 5. Agenda-specific Speech Analysis

Filter speeches by keyword to analyze specific policy topics.

```{r agenda-keywords, exercise=TRUE}
agendas <- list(
  "AI"            = "\uc778\uacf5\uc9c0\ub2a5|AI|\uc5d0\uc774\uc544\uc774",
  "Semiconductor" = "\ubc18\ub3c4\uccb4|\uce69|\ud30c\uc6b4\ub4dc\ub9ac",
  "Broadcasting"  = "\ubc29\uc1a1|\ubbf8\ub514\uc5b4|OTT|\ub137\ud50c\ub9ad\uc2a4",
  "Telecom"       = "\ud1b5\uc2e0|5G|\uc774\ub3d9\ud1b5\uc2e0|\uc8fc\ud30c\uc218",
  "Privacy"       = "\uac1c\uc778\uc815\ubcf4|\ud504\ub77c\uc774\ubc84\uc2dc|\ub370\uc774\ud130\ubcf4\ud638"
)

agenda_counts <- do.call(rbind, lapply(names(agendas), function(topic) {
  n <- sum(str_detect(speeches$speech, agendas[[topic]]))
  data.frame(topic = topic, n_mentions = n)
}))

ggplot(agenda_counts, aes(x = reorder(topic, n_mentions), y = n_mentions)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(x = NULL, y = "Related Speeches",
       title = "Speech Frequency by Agenda (22nd, Sci & ICT)") +
  theme_minimal()
```

### Agenda Trends Over Time

```{r agenda-trend, exercise=TRUE}
agendas <- list(
  "AI"            = "\uc778\uacf5\uc9c0\ub2a5|AI|\uc5d0\uc774\uc544\uc774",
  "Semiconductor" = "\ubc18\ub3c4\uccb4|\uce69|\ud30c\uc6b4\ub4dc\ub9ac",
  "Broadcasting"  = "\ubc29\uc1a1|\ubbf8\ub514\uc5b4|OTT|\ub137\ud50c\ub9ad\uc2a4",
  "Telecom"       = "\ud1b5\uc2e0|5G|\uc774\ub3d9\ud1b5\uc2e0|\uc8fc\ud30c\uc218",
  "Privacy"       = "\uac1c\uc778\uc815\ubcf4|\ud504\ub77c\uc774\ubc84\uc2dc|\ub370\uc774\ud130\ubcf4\ud638"
)

agenda_trend <- do.call(rbind, lapply(names(agendas), function(topic) {
  speeches %>%
    mutate(month = format(date, "%Y-%m")) %>%
    group_by(month) %>%
    summarise(
      mentions = sum(str_detect(speech, agendas[[topic]])),
      total = n(),
      .groups = "drop"
    ) %>%
    mutate(topic = topic, pct = mentions / total * 100)
}))

agenda_trend %>%
  mutate(month_date = as.Date(paste0(month, "-01"))) %>%
  ggplot(aes(x = month_date, y = pct, color = topic)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  labs(x = NULL, y = "Related Speech Proportion (%)",
       color = "Agenda",
       title = "Agenda Trends Over Time (22nd, Sci & ICT)") +
  theme_minimal()
```

Do any agendas spike at specific times? Try to identify what events might have caused the spikes.

## 6. Member Agenda Profiles

Even within the same committee, legislators focus on different topics.

```{r member-agenda, exercise=TRUE}
leg_speeches <- speeches %>%
  filter(role == "legislator")

agendas <- list(
  "AI"            = "\uc778\uacf5\uc9c0\ub2a5|AI|\uc5d0\uc774\uc544\uc774",
  "Semiconductor" = "\ubc18\ub3c4\uccb4|\uce69|\ud30c\uc6b4\ub4dc\ub9ac",
  "Broadcasting"  = "\ubc29\uc1a1|\ubbf8\ub514\uc5b4|OTT|\ub137\ud50c\ub9ad\uc2a4",
  "Telecom"       = "\ud1b5\uc2e0|5G|\uc774\ub3d9\ud1b5\uc2e0|\uc8fc\ud30c\uc218",
  "Privacy"       = "\uac1c\uc778\uc815\ubcf4|\ud504\ub77c\uc774\ubc84\uc2dc|\ub370\uc774\ud130\ubcf4\ud638"
)

member_agenda <- do.call(rbind, lapply(names(agendas), function(topic) {
  leg_speeches %>%
    group_by(speaker_name) %>%
    summarise(
      n_total = n(),
      n_topic = sum(str_detect(speech, agendas[[topic]])),
      .groups = "drop"
    ) %>%
    filter(n_total >= 20) %>%
    mutate(topic = topic, pct = n_topic / n_total * 100)
}))

member_agenda %>%
  group_by(topic) %>%
  slice_max(pct, n = 3) %>%
  select(topic, speaker_name, n_total, n_topic, pct) %>%
  arrange(topic, desc(pct))
```

### Member Agenda Heatmap

```{r member-heatmap, exercise=TRUE, fig.height=7}
leg_speeches <- speeches %>%
  filter(role == "legislator")

agendas <- list(
  "AI"            = "\uc778\uacf5\uc9c0\ub2a5|AI|\uc5d0\uc774\uc544\uc774",
  "Semiconductor" = "\ubc18\ub3c4\uccb4|\uce69|\ud30c\uc6b4\ub4dc\ub9ac",
  "Broadcasting"  = "\ubc29\uc1a1|\ubbf8\ub514\uc5b4|OTT|\ub137\ud50c\ub9ad\uc2a4",
  "Telecom"       = "\ud1b5\uc2e0|5G|\uc774\ub3d9\ud1b5\uc2e0|\uc8fc\ud30c\uc218",
  "Privacy"       = "\uac1c\uc778\uc815\ubcf4|\ud504\ub77c\uc774\ubc84\uc2dc|\ub370\uc774\ud130\ubcf4\ud638"
)

member_agenda <- do.call(rbind, lapply(names(agendas), function(topic) {
  leg_speeches %>%
    group_by(speaker_name) %>%
    summarise(
      n_total = n(),
      n_topic = sum(str_detect(speech, agendas[[topic]])),
      .groups = "drop"
    ) %>%
    filter(n_total >= 20) %>%
    mutate(topic = topic, pct = n_topic / n_total * 100)
}))

active_members <- leg_speeches %>%
  count(speaker_name) %>%
  filter(n >= 30) %>%
  pull(speaker_name)

heatmap_data <- member_agenda %>%
  filter(speaker_name %in% active_members) %>%
  select(speaker_name, topic, pct)

ggplot(heatmap_data, aes(x = topic, y = speaker_name, fill = pct)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(pct, 1)), size = 2.5) +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(x = NULL, y = NULL, fill = "Speech %",
       title = "Member Agenda Profile (22nd, Sci & ICT)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 7. Speech Diversity

Some legislators spread their attention across topics, while others concentrate on a few.

```{r diversity, exercise=TRUE}
leg_speeches <- speeches %>%
  filter(role == "legislator")

agendas <- list(
  "AI"            = "\uc778\uacf5\uc9c0\ub2a5|AI|\uc5d0\uc774\uc544\uc774",
  "Semiconductor" = "\ubc18\ub3c4\uccb4|\uce69|\ud30c\uc6b4\ub4dc\ub9ac",
  "Broadcasting"  = "\ubc29\uc1a1|\ubbf8\ub514\uc5b4|OTT|\ub137\ud50c\ub9ad\uc2a4",
  "Telecom"       = "\ud1b5\uc2e0|5G|\uc774\ub3d9\ud1b5\uc2e0|\uc8fc\ud30c\uc218",
  "Privacy"       = "\uac1c\uc778\uc815\ubcf4|\ud504\ub77c\uc774\ubc84\uc2dc|\ub370\uc774\ud130\ubcf4\ud638"
)

member_agenda <- do.call(rbind, lapply(names(agendas), function(topic) {
  leg_speeches %>%
    group_by(speaker_name) %>%
    summarise(
      n_total = n(),
      n_topic = sum(str_detect(speech, agendas[[topic]])),
      .groups = "drop"
    ) %>%
    filter(n_total >= 20) %>%
    mutate(topic = topic, pct = n_topic / n_total * 100)
}))

active_members <- leg_speeches %>%
  count(speaker_name) %>%
  filter(n >= 30) %>%
  pull(speaker_name)

member_concentration <- member_agenda %>%
  filter(speaker_name %in% active_members) %>%
  group_by(speaker_name) %>%
  summarise(
    max_pct = max(pct),
    n_topics_active = sum(pct > 5)
  )

ggplot(member_concentration, aes(x = n_topics_active, y = max_pct)) +
  geom_jitter(width = 0.2, alpha = 0.6, color = "steelblue", size = 3) +
  labs(
    x = "Active Agendas (>5% speeches)",
    y = "Max Agenda Share (%)",
    title = "Member Concentration vs Diversity in Agenda Coverage"
  ) +
  theme_minimal()
```

## Summary

| Analysis | What It Reveals |
|----------|-----------------|
| Speech frequency | Active vs passive members |
| Speech length | Question style (brief Q vs long statement) |
| Role comparison | Committee Q&A structure |
| Agenda tracking | Changes in policy attention over time |
| Member agenda profiles | Individual specialization |

> **Key insight**: Good descriptive statistics and visualization can reveal as much as complex models. Understanding your data thoroughly is always the first step.

## Exercises

### Exercise 1: Chair Speech Patterns

Analyze the chair's (`role == "chair"`) speech patterns. How does the chair's speech length compare to legislators and ministers?

```{r ex1, exercise=TRUE}
# Your code here

```

```{r ex1-solution}
speeches %>%
  mutate(speech_length = nchar(speech)) %>%
  filter(role %in% c("legislator", "chair", "minister")) %>%
  group_by(role) %>%
  summarise(
    n = n(),
    avg = round(mean(speech_length)),
    median = median(speech_length)
  )
```

### Exercise 2: Seniority and Speech Patterns

Join with `legislators` data using `member_id` and compare speech patterns by **seniority**. Do senior legislators speak more frequently and/or longer?

```{r ex2, exercise=TRUE}
# Your code here

```

```{r ex2-hint}
# Join leg_speeches with legislators (assembly == 22)
# Group by seniority
# Compare n_speeches per member and avg_length
leg_22 <- legislators %>%
  filter(assembly == 22) %>%
  select(member_id, seniority)

speeches %>%
  filter(role == "legislator") %>%
  mutate(speech_length = nchar(speech)) %>%
  inner_join(leg_22, by = "member_id") %>%
  group_by(seniority) %>%
  summarise(
    n_members = n_distinct(member_id),
    avg_speeches = round(n() / n_members),
    avg_length = round(mean(speech_length))
  )
```

### Exercise 3: Custom Keywords

Create your own keyword dictionary (e.g., "regulation", "innovation", "protection") and analyze usage patterns by member and over time.

```{r ex3, exercise=TRUE}
# Your code here
# Example:
# frames <- list(
#   "Regulation" = "...",
#   "Innovation" = "...",
#   "Protection" = "..."
# )

```

```{r ex3-hint}
frames <- list(
  "Regulation" = "\uaddc\uc81c|\uaddc\uc815|\uc81c\ud55c|\uae08\uc9c0",
  "Innovation" = "\ud601\uc2e0|\ubc1c\uc804|\uc131\uc7a5|\ucd09\uc9c4",
  "Protection" = "\ubcf4\ud638|\uc548\uc804|\ud53c\ud574|\ubc29\uc9c0"
)

# Apply the same pattern from the agenda analysis
```
