---
title: "Tutorial 3: Regression Analysis"
subtitle: "assemblykor - Korean National Assembly Data for Political Science Education"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(assemblykor)
library(dplyr)
library(ggplot2)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.width = 7, fig.height = 4)
data(legislators)
data(wealth)
data(seminars)
set_ko_font()

# Pre-build merged dataset
df <- legislators %>%
  filter(assembly == 22) %>%
  inner_join(wealth %>% filter(year == 2024), by = "member_id") %>%
  mutate(log_nw = log(pmax(net_worth, 1)))

df_3party <- df %>%
  filter(party %in% c("\uad6d\ubbfc\uc758\ud798", "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9", "\uc870\uad6d\ud601\uc2e0\ub2f9"))

tutorial_options(exercise.eval = FALSE)
```

```{r theme, echo=FALSE}
htmltools::includeCSS(
  system.file("tutorials/css/assemblykor.css", package = "assemblykor")
)
```

## Learning Objectives

**Research Question**: Does legislator wealth affect legislative activity?

You will learn:

- Running OLS regression with `lm()` and reading output
- The role of control variables (omitted variable bias)
- Log transformation for skewed variables
- Interaction terms
- Visualizing regression results (coefficient plot)

## 1. What Is Regression?

**Core idea**: "When X changes by one unit, how much does Y change on average?"

$$Y = \beta_0 + \beta_1 X + \epsilon$$

- $\beta_0$ (intercept): predicted Y when X = 0
- $\beta_1$ (slope): average change in Y per unit change in X
- $\epsilon$ (error): unexplained variation

## 2. Data Preparation

```{r prep, exercise=TRUE}
df <- legislators %>%
  filter(assembly == 22) %>%
  inner_join(wealth %>% filter(year == 2024), by = "member_id")

cat("Sample size:", nrow(df), "legislators\n")
```

### Visual: Scatter with Regression Line

```{r ols-visual, exercise=TRUE, fig.height=4}
ggplot(df, aes(x = net_worth / 1e6, y = n_bills_lead)) +
  geom_point(alpha = 0.4, color = "steelblue") +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  labs(x = "Net Worth (billion KRW)", y = "Lead-sponsored Bills",
       title = "Net Worth and Legislative Activity",
       subtitle = "Red line = regression line, gray = 95% CI") +
  theme_minimal()
```

## 3. Simple Regression

```{r simple-ols, exercise=TRUE}
m1 <- lm(n_bills_lead ~ net_worth, data = df)
summary(m1)
```

### Reading the Output

| Item | Location | Meaning |
|------|----------|---------|
| `Estimate` | Coefficients | Estimated $\hat{\beta}$ |
| `Std. Error` | Coefficients | Uncertainty of estimate |
| `Pr(>|t|)` | Coefficients | p-value (<0.05 = significant) |
| `R-squared` | Bottom | Proportion of variance explained |

> **About p-values**: A small p-value means the relationship is unlikely due to chance, NOT that the effect is large. Coefficient magnitude and statistical significance are separate things.

```{r quiz-interpret}
quiz(
  question("Why is the net_worth coefficient extremely small?",
    answer("Small sample size"),
    answer("The variable is measured in thousands of KRW, so numbers are very large", correct = TRUE),
    answer("Too many missing values"),
    answer("The dependent variable is not continuous"),
    correct = "Correct! Converting to billions makes interpretation easier."
  )
)
```

## 4. Log Transformation

Wealth data is typically right-skewed. Log transformation helps.

```{r log-model, exercise=TRUE}
df <- df %>%
  mutate(log_nw = log(pmax(net_worth, 1)))

m2 <- lm(n_bills_lead ~ log_nw, data = df)
summary(m2)
```

> **Interpretation**: In a log-linear model, when log(X) increases by 1 (X roughly triples), Y changes by $\beta$.

## 5. Multiple Regression: Adding Controls

Simple regression limitation: Is it wealth causing more bills, or do wealthier legislators also happen to be more senior?

**Multiple regression** estimates the **independent effect** of each variable.

```{r multiple, exercise=TRUE}
m3 <- lm(n_bills_lead ~ log_nw + gender + seniority + district_type, data = df)
summary(m3)
```

### Coefficient Comparison

```{r coef-compare, exercise=TRUE}
m2 <- lm(n_bills_lead ~ log_nw, data = df)
m3 <- lm(n_bills_lead ~ log_nw + gender + seniority + district_type, data = df)

data.frame(
  Model = c("Bivariate (M2)", "With Controls (M3)"),
  log_nw_coef = c(round(coef(m2)["log_nw"], 3), round(coef(m3)["log_nw"], 3)),
  R_squared = c(round(summary(m2)$r.squared, 3), round(summary(m3)$r.squared, 3))
)
```

If the coefficient shrinks after adding controls, part of the original effect was due to **confounding**.

```{r quiz-controls}
quiz(
  question("Why does adding control variables change the key coefficient?",
    answer("Controls create omitted variable bias (OVB) that biases the main coefficient", correct = TRUE),
    answer("Sample size decreases"),
    answer("Adding variables always shrinks coefficients"),
    answer("R-squared increases, so coefficients must change"),
    correct = "Correct! OVB occurs when omitted variables affect both X and Y."
  )
)
```

## 6. Categorical Variables

```{r categorical, exercise=TRUE}
df_3party <- df %>%
  filter(party %in% c("\uad6d\ubbfc\uc758\ud798", "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9", "\uc870\uad6d\ud601\uc2e0\ub2f9"))

m4 <- lm(n_bills_lead ~ log_nw + gender + seniority + district_type + party,
         data = df_3party)
summary(m4)
```

R automatically creates dummy variables. One category becomes the **reference**. Use `relevel()` to change it.

## 7. Interaction Terms

Does the wealth effect differ by gender?

```{r interaction, exercise=TRUE}
m5 <- lm(n_bills_lead ~ log_nw * gender + seniority + district_type, data = df)
summary(m5)
```

```{r interaction-plot, exercise=TRUE, fig.height=4}
ggplot(df, aes(x = log_nw, y = n_bills_lead, color = gender)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  scale_color_manual(values = c("F" = "#E66101", "M" = "#5E3C99"),
                     labels = c("Female", "Male")) +
  labs(x = "log(Net Worth)", y = "Lead Bills", color = "Gender",
       title = "Wealth and Bills: Gender Interaction") +
  theme_minimal()
```

Different slopes indicate an interaction effect.

## 8. Coefficient Plot

Visualizing regression results is more intuitive than reading tables.

```{r coef-plot, exercise=TRUE, fig.height=4}
m3 <- lm(n_bills_lead ~ log_nw + gender + seniority + district_type, data = df)

coef_df <- broom::tidy(m3, conf.int = TRUE) %>%
  filter(term != "(Intercept)")

ggplot(coef_df, aes(x = estimate, y = reorder(term, estimate))) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high),
                  color = "steelblue", size = 0.8) +
  labs(x = "Coefficient (95% CI)", y = NULL,
       title = "Determinants of Bill Lead-Sponsorship") +
  theme_minimal(base_size = 13)
```

- CI crossing 0 = not significant
- Left of 0 = negative effect, right = positive effect

## 9. Model Diagnostics

```{r diagnostics, exercise=TRUE, fig.height=6}
m3 <- lm(n_bills_lead ~ log_nw + gender + seniority + district_type, data = df)
par(mfrow = c(2, 2))
plot(m3)
```

| Plot | Checks | Good Sign |
|------|--------|-----------|
| Residuals vs Fitted | Homoscedasticity | No pattern |
| Q-Q Plot | Normality | Points on diagonal |
| Scale-Location | Homoscedasticity | Horizontal line |
| Residuals vs Leverage | Influential points | No extreme outliers |

## 10. Results Summary

```{r results-table, exercise=TRUE}
m1 <- lm(n_bills_lead ~ net_worth, data = df)
m2 <- lm(n_bills_lead ~ log_nw, data = df)
m3 <- lm(n_bills_lead ~ log_nw + gender + seniority + district_type, data = df)

models <- list(m1, m2, m3)
data.frame(
  Model = c("M1: Raw Scale", "M2: Log", "M3: + Controls"),
  Coef = sapply(models, \(m) round(coef(m)[2], 4)),
  p = sapply(models, \(m) round(summary(m)$coef[2, 4], 4)),
  R2 = sapply(models, \(m) round(summary(m)$r.squared, 3)),
  N = sapply(models, \(m) nobs(m))
)
```

## Exercises

### Exercise 1: Real Estate

Use `real_estate` as the predictor instead of `net_worth`. Compare results.

```{r ex1, exercise=TRUE}

```

```{r ex1-solution}
m_re <- lm(n_bills_lead ~ real_estate, data = df)
m_nw <- lm(n_bills_lead ~ net_worth, data = df)
summary(m_re)
```

### Exercise 2: Gangnam Property

Add `has_gangnam_property` as a control variable. Is it significant?

```{r ex2, exercise=TRUE}

```

```{r ex2-solution}
m_gangnam <- lm(n_bills_lead ~ log_nw + has_gangnam_property + gender +
                  seniority + district_type, data = df)
summary(m_gangnam)
```

### Exercise 3: Seminar Analysis with Coefficient Plot

Analyze whether seminar count (`n_seminars`) affects cross-party ratio. Control for governing status and seniority. Visualize with a coefficient plot.

```{r ex3, exercise=TRUE}

```

```{r ex3-solution}
data(seminars)
m_sem <- lm(cross_party_ratio ~ n_seminars + is_governing + seniority,
            data = seminars)

broom::tidy(m_sem, conf.int = TRUE) %>%
  filter(term != "(Intercept)") %>%
  ggplot(aes(x = estimate, y = reorder(term, estimate))) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high), color = "steelblue") +
  labs(x = "Coefficient (95% CI)", y = NULL,
       title = "Determinants of Cross-party Cooperation") +
  theme_minimal()
```

### Exercise 4: Log-Log Model

Estimate a log-log model and interpret the coefficient as **elasticity**.

```{r ex4, exercise=TRUE}

```

```{r ex4-solution}
df_pos <- df %>% filter(n_bills_lead > 0, net_worth > 0)
m_loglog <- lm(log(n_bills_lead) ~ log(net_worth), data = df_pos)
summary(m_loglog)
# Interpretation: 1% increase in net worth -> beta% change in bills
```
