---
title: "Tutorial 4: Panel Data and Fixed Effects"
subtitle: "assemblykor - Korean National Assembly Data for Political Science Education"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(assemblykor)
library(dplyr)
library(ggplot2)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.width = 7, fig.height = 4)
data(wealth)
data(seminars)
set_ko_font()
tutorial_options(exercise.eval = FALSE)
```

```{r theme, echo=FALSE}
htmltools::includeCSS(
  system.file("tutorials/css/assemblykor.css", package = "assemblykor")
)
```

## Learning Objectives

**Research Question**: Does becoming the governing party reduce cross-party cooperation?

You will learn:

- Panel data structure (unit x time)
- Between vs within variation
- Pooled OLS, individual FE, two-way FE
- `fixest::feols()` for fixed effects estimation
- Cluster standard errors

## 1. What Is Panel Data?

**Panel data** observes the same units repeatedly over time. The `wealth` data is a legislator-year panel.

```{r panel-structure, exercise=TRUE}
cat("Observations:", nrow(wealth), "\n")
cat("Legislators:", n_distinct(wealth$member_id), "\n")
cat("Year range:", paste(range(wealth$year), collapse = "-"), "\n")

obs_per_member <- wealth %>% count(member_id) %>% pull(n)
table(obs_per_member)
```

Not all legislators are observed the same number of times. This is an **unbalanced panel**.

```{r quiz-panel}
quiz(
  question("Which of the following is an example of panel data?",
    answer("Surveying 1,000 different people each year for 5 years"),
    answer("Tracking the same 200 legislators' wealth annually for 10 years", correct = TRUE),
    answer("Recording daily stock prices for a single company"),
    answer("Comparing GDP across 50 countries in 2024"),
    correct = "Correct! Panel data requires the SAME units observed repeatedly. A new random sample each year is repeated cross-section, not panel."
  )
)
```

## 2. Between vs Within Variation

- **Between**: Rich legislator vs poor legislator (level differences)
- **Within**: How does the SAME legislator's wealth change over time?

```{r between-within, exercise=TRUE, fig.height=4}
set.seed(42)
sample_ids <- wealth %>%
  count(member_id) %>%
  filter(n >= 6) %>%
  slice_sample(n = 5) %>%
  pull(member_id)

wealth %>%
  filter(member_id %in% sample_ids) %>%
  ggplot(aes(x = year, y = net_worth / 1e6, color = name, group = member_id)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(x = "Year", y = "Net Worth (billion KRW)", color = "Legislator",
       title = "Individual Net Worth Trajectories (5 legislators)") +
  theme_minimal()
```

**Fixed effects** uses ONLY within variation: "What changed within the same person?"

## 3. Pooled OLS

Ignoring individual differences:

```{r pooled, exercise=TRUE}
m_pooled <- lm(deposits ~ real_estate + n_properties, data = wealth)
summary(m_pooled)
```

**Problem**: Unobserved individual traits (family background, business ability) can bias estimates. This is **omitted variable bias (OVB)**.

## 4. Individual Fixed Effects

Each legislator gets their own intercept, absorbing all time-invariant personal characteristics.

```{r fe-individual, exercise=TRUE}
library(fixest)
m_fe <- feols(deposits ~ real_estate + n_properties | member_id, data = wealth)
summary(m_fe)
```

`| member_id` specifies the fixed effect. This estimates the effect of real estate on deposits **within the same legislator**.

### How It Works

1. Calculate each legislator's personal mean
2. Subtract the mean from all observations (demeaning)
3. Run OLS on the demeaned data

Result: between-person differences are completely eliminated.

```{r quiz-fe}
quiz(
  question("A legislator from a wealthy family always has higher deposits. With individual FE, this family background effect is:",
    answer("Estimated as a separate coefficient in the output"),
    answer("Absorbed into the individual intercept automatically", correct = TRUE),
    answer("Included as an additional observable control variable"),
    answer("Still causing bias in the real estate coefficient"),
    correct = "Correct! FE absorbs ALL time-invariant traits (family background, innate ability, etc.) into each person's intercept."
  )
)
```

### Pooled OLS vs Fixed Effects

```{r compare-pooled-fe, exercise=TRUE}
library(fixest)
m_pooled <- lm(deposits ~ real_estate + n_properties, data = wealth)
m_fe <- feols(deposits ~ real_estate + n_properties | member_id, data = wealth)

etable(m_pooled, m_fe, headers = c("Pooled OLS", "Individual FE"))
```

## 5. Two-Way Fixed Effects

Add year FE to control for common time trends (e.g., housing price booms).

```{r twfe, exercise=TRUE}
library(fixest)
m_pooled <- lm(deposits ~ real_estate + n_properties, data = wealth)
m_fe <- feols(deposits ~ real_estate + n_properties | member_id, data = wealth)
m_twfe <- feols(deposits ~ real_estate + n_properties | member_id + year, data = wealth)

etable(m_pooled, m_fe, m_twfe,
       headers = c("Pooled", "Individual FE", "Two-way FE"))
```

```{r quiz-twfe}
quiz(
  question("In 2021, housing prices surged nationwide, increasing all legislators' real estate values. Which model handles this?",
    answer("Pooled OLS"),
    answer("Individual FE only"),
    answer("Two-way FE (individual + year)", correct = TRUE),
    answer("None of the above"),
    correct = "Correct! Year FE absorbs common shocks like a nationwide housing boom that affects everyone equally."
  )
)
```

## 6. DiD with Seminars Data

**Research Question**: Does becoming the governing party reduce cross-party cooperation?

```{r did-prep, exercise=TRUE}
sem_panel <- seminars %>%
  filter(n_seminars > 0, year >= 2016)

cat("Observations:", nrow(sem_panel), "\n")
cat("Legislators:", n_distinct(sem_panel$name), "\n")

switchers <- sem_panel %>%
  group_by(name) %>%
  summarise(ever_gov = any(is_governing), ever_opp = any(!is_governing)) %>%
  summarise(n = sum(ever_gov & ever_opp))

cat("Switchers (experienced both):", switchers$n, "\n")
```

### DiD Visualization

```{r did-visual, exercise=TRUE, fig.height=4}
sem_panel <- seminars %>%
  filter(n_seminars > 0, year >= 2016)

sem_panel %>%
  group_by(year, is_governing) %>%
  summarise(mean_cross = mean(cross_party_ratio, na.rm = TRUE), .groups = "drop") %>%
  mutate(status = ifelse(is_governing, "Government", "Opposition")) %>%
  ggplot(aes(x = year, y = mean_cross, color = status)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  scale_color_manual(values = c("Government" = "tomato", "Opposition" = "dodgerblue")) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Year", y = "Cross-party Seminar Ratio", color = NULL,
       title = "Government vs Opposition Cross-party Cooperation") +
  theme_minimal()
```

### DiD Estimation

```{r did-estimate, exercise=TRUE}
library(fixest)
sem_panel <- seminars %>%
  filter(n_seminars > 0, year >= 2016)

m_did <- feols(cross_party_ratio ~ is_governing + n_seminars | name + year,
               data = sem_panel)
summary(m_did)
```

### Cluster Standard Errors

Panel data errors may be correlated within units. Cluster SE accounts for this.

```{r cluster-se, exercise=TRUE}
library(fixest)
sem_panel <- seminars %>%
  filter(n_seminars > 0, year >= 2016)

m_did <- feols(cross_party_ratio ~ is_governing + n_seminars | name + year,
               data = sem_panel)
summary(m_did, cluster = "name")
```

## 7. FE Limitations

### Time-invariant variables cannot be estimated

```{r time-invariant, exercise=TRUE}
library(fixest)
sem_panel <- seminars %>%
  filter(n_seminars > 0, year >= 2016)

m_gender <- feols(cross_party_ratio ~ is_governing + is_female | name + year,
                  data = sem_panel)
summary(m_gender)
```

`is_female` is absorbed by the individual FE because it never changes within a person.

## Summary

| Model | Controls For | Limitation |
|-------|-------------|-----------|
| Pooled OLS | Only included variables | OVB from unobservables |
| Individual FE | All time-invariant traits | Can't estimate time-invariant effects |
| Two-way FE | + Common time shocks | Can't estimate individual time trends |
| DiD (FE-based) | + Treatment effect | Requires parallel trends assumption |

## Exercises

### Exercise 1: Stocks and Real Estate

Estimate individual FE model: does real estate growth predict stock growth for the same legislator?

```{r ex1, exercise=TRUE}

```

```{r ex1-solution}
library(fixest)
m <- feols(stocks ~ real_estate | member_id, data = wealth)
summary(m)
```

### Exercise 2: Time-Invariant Problem

What happens when you use `has_seoul_property` with individual FE? Explain why.

```{r ex2, exercise=TRUE}

```

```{r ex2-hint}
# Check if has_seoul_property changes over time
wealth %>%
  group_by(member_id) %>%
  summarise(changes = n_distinct(has_seoul_property)) %>%
  count(changes)
```

### Exercise 3: Three-Model Comparison

Compare Pooled OLS, Individual FE, and Two-way FE using `etable()`. Which is most appropriate?

```{r ex3, exercise=TRUE}

```

```{r ex3-solution}
library(fixest)
m1 <- lm(deposits ~ real_estate + stocks, data = wealth)
m2 <- feols(deposits ~ real_estate + stocks | member_id, data = wealth)
m3 <- feols(deposits ~ real_estate + stocks | member_id + year, data = wealth)
etable(m1, m2, m3, headers = c("Pooled", "Individual FE", "Two-way FE"))
```
