---
title: "Tutorial 7: Roll Call Analysis"
subtitle: "assemblykor - Korean National Assembly Data for Political Science Education"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(assemblykor)
library(dplyr)
library(tidyr)
library(ggplot2)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
data(votes)
data(roll_calls)
data(legislators)
set_ko_font()
tutorial_options(exercise.eval = FALSE)
```

```{r theme, echo=FALSE}
htmltools::includeCSS(
  system.file("tutorials/css/assemblykor.css", package = "assemblykor")
)
```

## Learning Objectives

**Research Question**: How cohesive are Korean parties in legislative voting?

In a democracy, **roll call voting** is a core function of the legislature. Analyzing who votes yes and no reveals party cohesion, rebel members, and ideological distances between legislators.

You will learn:

- The structure of plenary vote data
- Measuring party cohesion: Rice Index
- Identifying rebel voters
- Visualizing voting patterns: heatmaps
- Measuring inter-legislator agreement

## 1. Understanding Vote Data

`assemblykor` includes two voting datasets:

| Dataset | Level | Scope | Rows |
|---------|-------|-------|------|
| `votes` | **bill-level** aggregate | 20th-22nd | 7,997 |
| `roll_calls` | **member-level** record | 22nd only | 368,210 |

### votes: Bill-level results

```{r votes-overview, exercise=TRUE}
cat("Number of bills:", nrow(votes), "\n")
cat("By assembly:\n")
table(votes$assembly)
```

```{r votes-example, exercise=TRUE}
votes %>%
  select(bill_name, assembly, voted, yes, no, abstain) %>%
  head(5)
```

```{r quiz-votes, echo=FALSE}
quiz(
  question("A bill has voted = 260, yes = 135, no = 120, abstain = 5. What can you infer?",
    answer("The bill clearly passed with a comfortable margin"),
    answer("This was a contentious vote, nearly split in half", correct = TRUE),
    answer("Most legislators chose to abstain from the vote"),
    answer("The data contains an error in the vote counts"),
    correct = "Correct! A near-even yes/no split signals a controversial bill. Note that yes + no + abstain = voted."
  )
)
```

### Yes Rate Distribution

Most bills pass with overwhelming majorities. Controversial bills are the minority.

```{r yes-rate-dist, exercise=TRUE}
votes <- votes %>%
  mutate(yes_rate = yes / voted)

ggplot(votes, aes(x = yes_rate)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "white") +
  scale_x_continuous(labels = scales::percent) +
  labs(
    x = "Yes Rate",
    y = "Number of Bills",
    title = "Distribution of Yes Rates in Plenary Votes (20th-22nd)"
  ) +
  theme_minimal()
```

### Finding Contentious Bills

```{r contentious, exercise=TRUE}
votes %>%
  mutate(yes_rate = yes / voted) %>%
  filter(yes_rate < 0.7) %>%
  arrange(yes_rate) %>%
  select(bill_name, assembly, yes_rate, yes, no) %>%
  head(10)
```

## 2. roll_calls: Member-level Records

```{r rc-overview, exercise=TRUE}
cat("Vote records:", nrow(roll_calls), "\n")
cat("Members:", n_distinct(roll_calls$member_id), "\n")
cat("Bills:", n_distinct(roll_calls$bill_id), "\n")

# Vote types
table(roll_calls$vote)
```

```{r rc-recode, exercise=TRUE}
# Recode votes to English for analysis convenience
roll_calls <- roll_calls %>%
  mutate(vote_en = case_match(vote,
    "\ucc2c\uc131" ~ "yes",
    "\ubc18\ub300" ~ "no",
    "\uae30\uad8c" ~ "abstain",
    .default = "absent"
  ))

table(roll_calls$vote_en)
```

## 3. Party Cohesion: Rice Index

The **Rice Index** is a classic measure of party cohesion:

$$\text{Rice Index} = |\text{proportion yes} - \text{proportion no}|$$

- 1: all members voted the same way (perfect cohesion)
- 0: evenly split between yes and no (complete division)

### Calculate Rice Index by Bill

```{r rice-calc, exercise=TRUE}
# Only actual votes (exclude absences)
rc_voted <- roll_calls %>%
  mutate(vote_en = case_match(vote,
    "\ucc2c\uc131" ~ "yes",
    "\ubc18\ub300" ~ "no",
    "\uae30\uad8c" ~ "abstain",
    .default = "absent"
  )) %>%
  filter(vote_en %in% c("yes", "no"))

# Rice Index by bill and party
rice_by_bill <- rc_voted %>%
  group_by(bill_id, party) %>%
  summarise(
    n_voted = n(),
    n_yes = sum(vote_en == "yes"),
    n_no = sum(vote_en == "no"),
    rice = abs(n_yes - n_no) / n_voted,
    .groups = "drop"
  ) %>%
  filter(n_voted >= 5)

# Major parties' average Rice Index
rice_by_bill %>%
  filter(party %in% c("\uad6d\ubbfc\uc758\ud798", "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9", "\uc870\uad6d\ud601\uc2e0\ub2f9")) %>%
  group_by(party) %>%
  summarise(
    n_votes = n(),
    mean_rice = round(mean(rice), 3),
    median_rice = round(median(rice), 3),
    min_rice = round(min(rice), 3)
  )
```

### Rice Index Distribution

```{r rice-plot, exercise=TRUE}
rc_voted <- roll_calls %>%
  mutate(vote_en = case_match(vote,
    "\ucc2c\uc131" ~ "yes",
    "\ubc18\ub300" ~ "no",
    "\uae30\uad8c" ~ "abstain",
    .default = "absent"
  )) %>%
  filter(vote_en %in% c("yes", "no"))

rice_by_bill <- rc_voted %>%
  group_by(bill_id, party) %>%
  summarise(
    n_voted = n(),
    n_yes = sum(vote_en == "yes"),
    n_no = sum(vote_en == "no"),
    rice = abs(n_yes - n_no) / n_voted,
    .groups = "drop"
  ) %>%
  filter(n_voted >= 5)

rice_by_bill %>%
  filter(party %in% c("\uad6d\ubbfc\uc758\ud798", "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9", "\uc870\uad6d\ud601\uc2e0\ub2f9")) %>%
  ggplot(aes(x = rice, fill = party)) +
  geom_histogram(bins = 30, alpha = 0.7, position = "identity") +
  scale_fill_manual(values = c(
    "\uad6d\ubbfc\uc758\ud798" = "#E61E2B",
    "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9" = "#004EA2",
    "\uc870\uad6d\ud601\uc2e0\ub2f9" = "#00B0F0"
  )) +
  facet_wrap(~ party, ncol = 1) +
  labs(
    x = "Rice Index", y = "Number of Bills",
    fill = NULL,
    title = "Party Cohesion (Rice Index) Distribution"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r quiz-rice, echo=FALSE}
quiz(
  question("Party A: 90 yes, 10 no. Party B: 55 yes, 45 no. Which party has higher Rice Index?",
    answer("Party B, because more members voted"),
    answer("Party A, because |0.9 - 0.1| = 0.8 > |0.55 - 0.45| = 0.1", correct = TRUE),
    answer("They are equal because both have a majority voting yes"),
    answer("Cannot compare without knowing the bill topic"),
    correct = "Correct! Party A (Rice = 0.8) is far more cohesive than Party B (Rice = 0.1), where the vote was nearly split."
  )
)
```

## 4. Finding Rebel Voters

A **rebel vote** occurs when a member votes against the party majority.

```{r rebel-calc, exercise=TRUE}
rc_voted <- roll_calls %>%
  mutate(vote_en = case_match(vote,
    "\ucc2c\uc131" ~ "yes",
    "\ubc18\ub300" ~ "no",
    "\uae30\uad8c" ~ "abstain",
    .default = "absent"
  )) %>%
  filter(vote_en %in% c("yes", "no"))

# Determine party majority per bill
party_majority <- rc_voted %>%
  group_by(bill_id, party) %>%
  summarise(
    majority_vote = ifelse(sum(vote_en == "yes") >= sum(vote_en == "no"),
                           "yes", "no"),
    .groups = "drop"
  )

# Identify rebel votes
rebel_votes <- rc_voted %>%
  inner_join(party_majority, by = c("bill_id", "party")) %>%
  mutate(is_rebel = vote_en != majority_vote)

# Rebel rate per member
rebel_rate <- rebel_votes %>%
  group_by(member_id, member_name, party) %>%
  summarise(
    n_votes = n(),
    n_rebel = sum(is_rebel),
    rebel_rate = round(n_rebel / n_votes, 3),
    .groups = "drop"
  ) %>%
  filter(n_votes >= 50)

rebel_rate %>%
  filter(party %in% c("\uad6d\ubbfc\uc758\ud798", "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9")) %>%
  arrange(desc(rebel_rate)) %>%
  head(10)
```

### Rebel Vote Distribution by Party

```{r rebel-plot, exercise=TRUE}
rc_voted <- roll_calls %>%
  mutate(vote_en = case_match(vote,
    "\ucc2c\uc131" ~ "yes",
    "\ubc18\ub300" ~ "no",
    "\uae30\uad8c" ~ "abstain",
    .default = "absent"
  )) %>%
  filter(vote_en %in% c("yes", "no"))

party_majority <- rc_voted %>%
  group_by(bill_id, party) %>%
  summarise(
    majority_vote = ifelse(sum(vote_en == "yes") >= sum(vote_en == "no"),
                           "yes", "no"),
    .groups = "drop"
  )

rebel_votes <- rc_voted %>%
  inner_join(party_majority, by = c("bill_id", "party")) %>%
  mutate(is_rebel = vote_en != majority_vote)

rebel_rate <- rebel_votes %>%
  group_by(member_id, member_name, party) %>%
  summarise(
    n_votes = n(),
    n_rebel = sum(is_rebel),
    rebel_rate = round(n_rebel / n_votes, 3),
    .groups = "drop"
  ) %>%
  filter(n_votes >= 50)

rebel_rate %>%
  filter(party %in% c("\uad6d\ubbfc\uc758\ud798", "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9", "\uc870\uad6d\ud601\uc2e0\ub2f9")) %>%
  ggplot(aes(x = party, y = rebel_rate, fill = party)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  scale_fill_manual(values = c(
    "\uad6d\ubbfc\uc758\ud798" = "#E61E2B",
    "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9" = "#004EA2",
    "\uc870\uad6d\ud601\uc2e0\ub2f9" = "#00B0F0"
  )) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = NULL, y = "Rebel Vote Rate",
    title = "Rebel Vote Rates by Party (22nd Assembly)"
  ) +
  theme_minimal()
```

## 5. Voting Pattern Heatmap

Visualizing voting patterns reveals the party structure at a glance.

```{r heatmap, exercise=TRUE, fig.height=8}
votes_with_rate <- votes %>%
  mutate(yes_rate = yes / voted)

# Contentious bills: 50-85% yes rate
contentious_bills <- votes_with_rate %>%
  filter(assembly == 22, yes_rate >= 0.45, yes_rate <= 0.85) %>%
  arrange(yes_rate) %>%
  head(30) %>%
  pull(bill_id)

# Member-level votes for contentious bills (two major parties)
rc_heat <- roll_calls %>%
  mutate(vote_en = case_match(vote,
    "\ucc2c\uc131" ~ "yes",
    "\ubc18\ub300" ~ "no",
    "\uae30\uad8c" ~ "abstain",
    .default = "absent"
  )) %>%
  filter(bill_id %in% contentious_bills,
         party %in% c("\uad6d\ubbfc\uc758\ud798", "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9")) %>%
  mutate(vote_num = case_match(vote_en,
    "yes" ~ 1, "no" ~ -1, "abstain" ~ 0, .default = NA_real_
  )) %>%
  filter(!is.na(vote_num))

# Vote matrix (member x bill)
vote_matrix <- rc_heat %>%
  select(member_name, bill_id, vote_num, party) %>%
  pivot_wider(names_from = bill_id, values_from = vote_num, values_fill = 0) %>%
  arrange(party, member_name)

# Convert back to long format for ggplot
heat_data <- vote_matrix %>%
  pivot_longer(cols = -c(member_name, party),
               names_to = "bill_id", values_to = "vote_num") %>%
  mutate(
    vote_label = case_match(vote_num,
      1 ~ "Yes", -1 ~ "No", 0 ~ "Abstain/Absent"
    ),
    member_name = factor(member_name, levels = vote_matrix$member_name)
  )

ggplot(heat_data, aes(x = bill_id, y = member_name, fill = vote_label)) +
  geom_tile(color = "white", linewidth = 0.1) +
  scale_fill_manual(values = c(
    "Yes" = "#4DAF4A", "No" = "#E41A1C", "Abstain/Absent" = "#CCCCCC"
  )) +
  labs(
    x = "Bill", y = NULL, fill = "Vote",
    title = "Member Voting Patterns on Contentious Bills (22nd)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  )
```

When colors are clearly divided by party (one side green, other side red), it indicates strong partisan voting.

## 6. Inter-member Voting Agreement

We can measure how similarly two legislators vote using **agreement rate**: the proportion of bills where they voted the same way.

```{r agreement, exercise=TRUE}
votes_with_rate <- votes %>%
  mutate(yes_rate = yes / voted)

contentious_bills <- votes_with_rate %>%
  filter(assembly == 22, yes_rate >= 0.45, yes_rate <= 0.85) %>%
  arrange(yes_rate) %>%
  head(30) %>%
  pull(bill_id)

rc_heat <- roll_calls %>%
  mutate(vote_en = case_match(vote,
    "\ucc2c\uc131" ~ "yes",
    "\ubc18\ub300" ~ "no",
    "\uae30\uad8c" ~ "abstain",
    .default = "absent"
  )) %>%
  filter(bill_id %in% contentious_bills,
         party %in% c("\uad6d\ubbfc\uc758\ud798", "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9")) %>%
  mutate(vote_num = case_match(vote_en,
    "yes" ~ 1, "no" ~ -1, "abstain" ~ 0, .default = NA_real_
  )) %>%
  filter(!is.na(vote_num))

vote_matrix <- rc_heat %>%
  select(member_name, bill_id, vote_num, party) %>%
  pivot_wider(names_from = bill_id, values_from = vote_num, values_fill = 0) %>%
  arrange(party, member_name)

# Sample 20 members (10 per party)
set.seed(42)
sample_names <- vote_matrix %>%
  group_by(party) %>%
  slice_sample(n = 10) %>%
  pull(member_name)

vote_mat <- vote_matrix %>%
  filter(member_name %in% sample_names) %>%
  select(-party)
member_names_ordered <- vote_mat$member_name
vote_mat <- vote_mat %>% select(-member_name) %>% as.matrix()
rownames(vote_mat) <- member_names_ordered

# Agreement matrix
n_bills_mat <- ncol(vote_mat)
agreement <- matrix(0, nrow(vote_mat), nrow(vote_mat))
for (i in 1:nrow(vote_mat)) {
  for (j in 1:nrow(vote_mat)) {
    agreement[i, j] <- sum(vote_mat[i, ] == vote_mat[j, ]) / n_bills_mat
  }
}
rownames(agreement) <- rownames(vote_mat)
colnames(agreement) <- rownames(vote_mat)

agree_df <- as.data.frame(as.table(agreement)) %>%
  rename(member1 = Var1, member2 = Var2, agreement = Freq)

ggplot(agree_df, aes(x = member1, y = member2, fill = agreement)) +
  geom_tile() +
  scale_fill_gradient2(
    low = "#E41A1C", mid = "white", high = "#4DAF4A",
    midpoint = 0.5, labels = scales::percent
  ) +
  labs(
    x = NULL, y = NULL, fill = "Agreement",
    title = "Inter-member Voting Agreement (22nd, Contentious Bills)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, size = 7),
    axis.text.y = element_text(size = 7)
  )
```

High agreement within party (green blocks) and low agreement across parties (red blocks) indicates that party is the key determinant of voting behavior.

## Summary

| Concept | Meaning | Measurement |
|---------|---------|-------------|
| Yes Rate | Bill controversy | `yes / voted` |
| Rice Index | Party cohesion | `|yes_prop - no_prop|` |
| Rebel Vote | Voting against party | Disagreement with majority |
| Agreement | Behavioral similarity | Pairwise agreement matrix |

These analyses form the foundation of **legislative behavior** research.

## Exercises

### Exercise 1: Absenteeism

Find the 10 members with the highest absence rate in the 22nd Assembly.

```{r ex1, exercise=TRUE}
# Your code here

```

```{r ex1-hint}
# Group by member, count absent votes, calculate rate
roll_calls %>%
  mutate(vote_en = case_match(vote,
    "\ucc2c\uc131" ~ "yes", "\ubc18\ub300" ~ "no",
    "\uae30\uad8c" ~ "abstain", .default = "absent")) %>%
  group_by(member_name, party) %>%
  summarise(n_total = n(), n_absent = sum(vote_en == "absent"),
            absent_rate = round(n_absent / n_total, 3), .groups = "drop") %>%
  arrange(desc(absent_rate)) %>%
  head(10)
```

### Exercise 2: Rice by Controversy

Compare Rice Index for contentious bills (yes rate 50-85%) vs non-contentious bills (90%+). What differences emerge?

```{r ex2, exercise=TRUE}
# Your code here

```

```{r ex2-hint}
# Join rice_by_bill with bill-level yes_rate, then compare
# Contentious bills naturally have lower Rice Index
# because parties split on controversial issues
```

### Exercise 3: Committee Cohesion

Compare party cohesion across different committees. Do some policy areas show more intra-party disagreement?

```{r ex3, exercise=TRUE}
# Your code here

```

```{r ex3-hint}
# Join rice_by_bill with votes to get committee info
# Group by party and committee, then compare mean Rice Index
```
