---
title: "Tutorial 6: Co-sponsorship Network Analysis"
subtitle: "assemblykor - Korean National Assembly Data for Political Science Education"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(assemblykor)
library(dplyr)
library(ggplot2)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.width = 7, fig.height = 6)
data(legislators)
data(bills)
set_ko_font()
tutorial_options(exercise.eval = FALSE)
```

```{r theme, echo=FALSE}
htmltools::includeCSS(
  system.file("tutorials/css/assemblykor.css", package = "assemblykor")
)
```

## Learning Objectives

**Research Question**: Who co-sponsors bills with whom?

Legislators do not propose bills alone. Beyond the lead sponsor, multiple legislators participate as co-sponsors. By constructing a **network (graph)** from these co-sponsorship ties, we can uncover cooperation structures, inter-party relationships, and key figures.

You will learn:

- What is a network: nodes and edges
- Building a network from co-sponsorship data
- Centrality measures: who is at the center?
- Community detection: finding tightly connected groups
- Network visualization and cross-party analysis

## 1. Network Basics

A network (graph) consists of two elements:

- **Node (vertex)**: unit of analysis. Here, **legislators**
- **Edge (link)**: relationship between nodes. Here, **co-sponsorship ties**

```{r concept, echo=FALSE, fig.height=3.5}
library(igraph)
g_ex <- graph_from_data_frame(
  data.frame(from = c("A", "A", "B", "C", "D"),
             to   = c("B", "C", "C", "D", "E")),
  directed = FALSE
)
set.seed(1)
plot(g_ex, vertex.size = 30, vertex.color = "lightblue",
     vertex.label.cex = 1.2, edge.width = 2,
     main = "Basic Network Structure")
```

In this figure:

- A, B, C, D, E are **nodes** (5)
- The lines are **edges** (5)
- C has 3 edges, giving it the highest **degree centrality**

## 2. Downloading Co-sponsorship Data

```{r download, exercise=TRUE, exercise.eval=FALSE}
proposers <- get_proposers()
dim(proposers)
head(proposers)
```

## 3. Building an Edge List

To create a network, we need an **edge list**: "who is connected to whom."

For each bill, we link the lead sponsor to each co-sponsor.

```{r edgelist, exercise=TRUE, exercise.eval=FALSE}
proposers <- get_proposers()

# Restrict to 22nd Assembly bills
props_22 <- proposers %>%
  filter(propose_date >= as.Date("2024-05-30"))

# Separate lead sponsors and co-sponsors
leads <- props_22 %>% filter(is_lead) %>% select(bill_id, lead = member_id)
cosponsors <- props_22 %>% filter(!is_lead) %>% select(bill_id, cosponsor = member_id)

# Create lead-cosponsor pairs
edges <- inner_join(leads, cosponsors, by = "bill_id") %>%
  filter(!is.na(lead), !is.na(cosponsor), lead != cosponsor)

# Edge weight: number of co-sponsored bills
edge_weights <- edges %>%
  count(lead, cosponsor, name = "weight")

cat("Nodes (legislators):", n_distinct(c(edge_weights$lead, edge_weights$cosponsor)), "\n")
cat("Edges (connections):", nrow(edge_weights), "\n")
cat("Mean weight:", round(mean(edge_weights$weight), 1), "\n")
```

```{r quiz-network}
quiz(
  question("What does an 'edge' represent in a co-sponsorship network?",
    answer("An individual legislator"),
    answer("A co-sponsorship tie between two legislators", correct = TRUE),
    answer("A single bill"),
    answer("Committee membership"),
    correct = "Correct! An edge represents the co-sponsorship connection between two legislators."
  )
)
```

## 4. Creating an igraph Network

```{r create-graph, exercise=TRUE, exercise.eval=FALSE}
library(igraph)
proposers <- get_proposers()

props_22 <- proposers %>%
  filter(propose_date >= as.Date("2024-05-30"))

leads <- props_22 %>% filter(is_lead) %>% select(bill_id, lead = member_id)
cosponsors <- props_22 %>% filter(!is_lead) %>% select(bill_id, cosponsor = member_id)

edge_weights <- inner_join(leads, cosponsors, by = "bill_id") %>%
  filter(!is.na(lead), !is.na(cosponsor), lead != cosponsor) %>%
  count(lead, cosponsor, name = "weight")

g <- graph_from_data_frame(edge_weights, directed = FALSE)

# Add legislator metadata
leg_22 <- legislators %>%
  filter(assembly == 22) %>%
  distinct(member_id, .keep_all = TRUE)

V(g)$party  <- leg_22$party[match(V(g)$name, leg_22$member_id)]
V(g)$gender <- leg_22$gender[match(V(g)$name, leg_22$member_id)]
V(g)$label  <- leg_22$name[match(V(g)$name, leg_22$member_id)]

cat("Nodes:", vcount(g), "\n")
cat("Edges:", ecount(g), "\n")
cat("Density:", round(graph.density(g), 4), "\n")
```

> **Network density**: the proportion of all possible connections that actually exist. Density = 1 means all nodes are connected; 0 means no connections. Co-sponsorship networks typically have quite high density.

## 5. Centrality Measures: Who Is at the Center?

There are multiple ways to define "important nodes" in a network. Each measure captures a different aspect of "importance."

### Degree Centrality

The simplest: **how many other legislators are you connected to?**

```{r degree, exercise=TRUE, exercise.eval=FALSE}
library(igraph)
proposers <- get_proposers()

props_22 <- proposers %>% filter(propose_date >= as.Date("2024-05-30"))
leads <- props_22 %>% filter(is_lead) %>% select(bill_id, lead = member_id)
cosponsors <- props_22 %>% filter(!is_lead) %>% select(bill_id, cosponsor = member_id)
edge_weights <- inner_join(leads, cosponsors, by = "bill_id") %>%
  filter(!is.na(lead), !is.na(cosponsor), lead != cosponsor) %>%
  count(lead, cosponsor, name = "weight")

g <- graph_from_data_frame(edge_weights, directed = FALSE)
leg_22 <- legislators %>% filter(assembly == 22) %>% distinct(member_id, .keep_all = TRUE)
V(g)$party <- leg_22$party[match(V(g)$name, leg_22$member_id)]
V(g)$label <- leg_22$name[match(V(g)$name, leg_22$member_id)]

degree_df <- data.frame(
  name = V(g)$label,
  party = V(g)$party,
  degree = degree(g),
  stringsAsFactors = FALSE
) %>% filter(!is.na(name))

degree_df %>% arrange(desc(degree)) %>% head(10)
```

### Betweenness Centrality

**Who serves as a "bridge" connecting different groups?** Betweenness measures the proportion of shortest paths passing through a node.

```{r betweenness, exercise=TRUE, exercise.eval=FALSE}
library(igraph)
proposers <- get_proposers()

props_22 <- proposers %>% filter(propose_date >= as.Date("2024-05-30"))
leads <- props_22 %>% filter(is_lead) %>% select(bill_id, lead = member_id)
cosponsors <- props_22 %>% filter(!is_lead) %>% select(bill_id, cosponsor = member_id)
edge_weights <- inner_join(leads, cosponsors, by = "bill_id") %>%
  filter(!is.na(lead), !is.na(cosponsor), lead != cosponsor) %>%
  count(lead, cosponsor, name = "weight")

g <- graph_from_data_frame(edge_weights, directed = FALSE)
leg_22 <- legislators %>% filter(assembly == 22) %>% distinct(member_id, .keep_all = TRUE)
V(g)$party <- leg_22$party[match(V(g)$name, leg_22$member_id)]
V(g)$label <- leg_22$name[match(V(g)$name, leg_22$member_id)]

between_df <- data.frame(
  name = V(g)$label,
  party = V(g)$party,
  betweenness = round(betweenness(g, normalized = TRUE), 6),
  stringsAsFactors = FALSE
) %>% filter(!is.na(name))

between_df %>% arrange(desc(betweenness)) %>% head(10)
```

Those with high degree and high betweenness may differ. High-betweenness legislators are not "people who know many people" but "people who connect different groups."

```{r quiz-centrality}
quiz(
  question("What role does a legislator with high betweenness centrality play?",
    answer("The legislator who proposed the most bills"),
    answer("A 'bridge' connecting different groups in the network", correct = TRUE),
    answer("A legislator active only within their own party"),
    answer("The most recently elected legislator"),
    correct = "Correct! High betweenness means connecting diverse groups within the network."
  )
)
```

### Centrality Comparison Scatter Plot

```{r centrality-scatter, exercise=TRUE, exercise.eval=FALSE, fig.height=5}
library(igraph)
proposers <- get_proposers()

props_22 <- proposers %>% filter(propose_date >= as.Date("2024-05-30"))
leads <- props_22 %>% filter(is_lead) %>% select(bill_id, lead = member_id)
cosponsors <- props_22 %>% filter(!is_lead) %>% select(bill_id, cosponsor = member_id)
edge_weights <- inner_join(leads, cosponsors, by = "bill_id") %>%
  filter(!is.na(lead), !is.na(cosponsor), lead != cosponsor) %>%
  count(lead, cosponsor, name = "weight")

g <- graph_from_data_frame(edge_weights, directed = FALSE)
leg_22 <- legislators %>% filter(assembly == 22) %>% distinct(member_id, .keep_all = TRUE)
V(g)$party <- leg_22$party[match(V(g)$name, leg_22$member_id)]
V(g)$label <- leg_22$name[match(V(g)$name, leg_22$member_id)]

centrality_df <- data.frame(
  name = V(g)$label,
  party = V(g)$party,
  degree = degree(g),
  betweenness = betweenness(g, normalized = TRUE),
  stringsAsFactors = FALSE
) %>% filter(!is.na(name), !is.na(party))

ggplot(centrality_df, aes(x = degree, y = betweenness)) +
  geom_point(alpha = 0.5, color = "steelblue") +
  labs(x = "Degree Centrality", y = "Betweenness Centrality",
       title = "Comparing Two Centrality Measures") +
  theme_minimal()
```

## 6. Party Network Structure

```{r party-stats, exercise=TRUE, exercise.eval=FALSE}
library(igraph)
proposers <- get_proposers()

props_22 <- proposers %>% filter(propose_date >= as.Date("2024-05-30"))
leads <- props_22 %>% filter(is_lead) %>% select(bill_id, lead = member_id)
cosponsors <- props_22 %>% filter(!is_lead) %>% select(bill_id, cosponsor = member_id)
edge_weights <- inner_join(leads, cosponsors, by = "bill_id") %>%
  filter(!is.na(lead), !is.na(cosponsor), lead != cosponsor) %>%
  count(lead, cosponsor, name = "weight")

g <- graph_from_data_frame(edge_weights, directed = FALSE)
leg_22 <- legislators %>% filter(assembly == 22) %>% distinct(member_id, .keep_all = TRUE)
V(g)$party <- leg_22$party[match(V(g)$name, leg_22$member_id)]
V(g)$label <- leg_22$name[match(V(g)$name, leg_22$member_id)]

centrality_df <- data.frame(
  name = V(g)$label,
  party = V(g)$party,
  degree = degree(g),
  betweenness = round(betweenness(g, normalized = TRUE), 6),
  stringsAsFactors = FALSE
) %>% filter(!is.na(name))

centrality_df %>%
  filter(party %in% c("\uad6d\ubbfc\uc758\ud798", "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9", "\uc870\uad6d\ud601\uc2e0\ub2f9")) %>%
  group_by(party) %>%
  summarise(
    n = n(),
    avg_degree = round(mean(degree), 1),
    max_degree = max(degree),
    avg_betweenness = round(mean(betweenness), 6)
  )
```

### Party Degree Distribution

```{r party-boxplot, exercise=TRUE, exercise.eval=FALSE, fig.height=4}
library(igraph)
proposers <- get_proposers()

props_22 <- proposers %>% filter(propose_date >= as.Date("2024-05-30"))
leads <- props_22 %>% filter(is_lead) %>% select(bill_id, lead = member_id)
cosponsors <- props_22 %>% filter(!is_lead) %>% select(bill_id, cosponsor = member_id)
edge_weights <- inner_join(leads, cosponsors, by = "bill_id") %>%
  filter(!is.na(lead), !is.na(cosponsor), lead != cosponsor) %>%
  count(lead, cosponsor, name = "weight")

g <- graph_from_data_frame(edge_weights, directed = FALSE)
leg_22 <- legislators %>% filter(assembly == 22) %>% distinct(member_id, .keep_all = TRUE)
V(g)$party <- leg_22$party[match(V(g)$name, leg_22$member_id)]

centrality_df <- data.frame(
  party = V(g)$party,
  degree = degree(g),
  stringsAsFactors = FALSE
) %>% filter(!is.na(party))

centrality_df %>%
  filter(party %in% c("\uad6d\ubbfc\uc758\ud798", "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9", "\uc870\uad6d\ud601\uc2e0\ub2f9")) %>%
  ggplot(aes(x = party, y = degree, fill = party)) +
  geom_boxplot(show.legend = FALSE, alpha = 0.7) +
  scale_fill_manual(values = c(
    "\uad6d\ubbfc\uc758\ud798" = "#E61E2B",
    "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9" = "#004EA2",
    "\uc870\uad6d\ud601\uc2e0\ub2f9" = "#00B0F0"
  )) +
  labs(x = NULL, y = "Degree Centrality",
       title = "Co-sponsorship Network Degree by Party") +
  theme_minimal()
```

## 7. Community Detection

Tightly connected subgroups within a network are called **communities**. Algorithms automatically find them.

### Louvain Algorithm

```{r community, exercise=TRUE, exercise.eval=FALSE}
library(igraph)
proposers <- get_proposers()

props_22 <- proposers %>% filter(propose_date >= as.Date("2024-05-30"))
leads <- props_22 %>% filter(is_lead) %>% select(bill_id, lead = member_id)
cosponsors <- props_22 %>% filter(!is_lead) %>% select(bill_id, cosponsor = member_id)
edge_weights <- inner_join(leads, cosponsors, by = "bill_id") %>%
  filter(!is.na(lead), !is.na(cosponsor), lead != cosponsor) %>%
  count(lead, cosponsor, name = "weight")

g <- graph_from_data_frame(edge_weights, directed = FALSE)
leg_22 <- legislators %>% filter(assembly == 22) %>% distinct(member_id, .keep_all = TRUE)
V(g)$party <- leg_22$party[match(V(g)$name, leg_22$member_id)]

comm <- cluster_louvain(g)

cat("Communities found:", max(membership(comm)), "\n")
cat("Modularity:", round(modularity(comm), 3), "\n")

# Community-party correspondence
comm_party <- data.frame(
  community = membership(comm),
  party = V(g)$party
) %>% filter(!is.na(party))

table(comm_party$community, comm_party$party) %>%
  as.data.frame() %>%
  filter(Freq > 3) %>%
  arrange(Var1, desc(Freq))
```

> **Modularity**: measures how clearly communities are separated. Close to 0 means no different from random; close to 1 means very distinct communities. Values above 0.3 typically indicate meaningful structure.

```{r quiz-community}
quiz(
  question("If the modularity value is 0.5, what does it mean?",
    answer("The network is divided into 5 communities"),
    answer("Within-community connections are much denser than in a random network", correct = TRUE),
    answer("50% of all legislators belong to one community"),
    answer("Cross-party connections are 50%"),
    correct = "Correct! Higher modularity means clearer community structure (close to 0 = random)."
  )
)
```

## 8. Cross-Party Connections

How much co-sponsorship happens across party lines?

```{r cross-party, exercise=TRUE, exercise.eval=FALSE}
library(igraph)
proposers <- get_proposers()

props_22 <- proposers %>% filter(propose_date >= as.Date("2024-05-30"))
leads <- props_22 %>% filter(is_lead) %>% select(bill_id, lead = member_id)
cosponsors <- props_22 %>% filter(!is_lead) %>% select(bill_id, cosponsor = member_id)
edge_weights <- inner_join(leads, cosponsors, by = "bill_id") %>%
  filter(!is.na(lead), !is.na(cosponsor), lead != cosponsor) %>%
  count(lead, cosponsor, name = "weight")

g <- graph_from_data_frame(edge_weights, directed = FALSE)
leg_22 <- legislators %>% filter(assembly == 22) %>% distinct(member_id, .keep_all = TRUE)
V(g)$party <- leg_22$party[match(V(g)$name, leg_22$member_id)]

edge_df <- as_data_frame(g, what = "edges") %>%
  mutate(
    party_from = V(g)$party[match(from, V(g)$name)],
    party_to   = V(g)$party[match(to, V(g)$name)],
    cross_party = party_from != party_to
  ) %>%
  filter(!is.na(party_from), !is.na(party_to))

cat("Total edges:", nrow(edge_df), "\n")
cat("Cross-party edges:", sum(edge_df$cross_party), "\n")
cat("Cross-party ratio:", round(mean(edge_df$cross_party) * 100, 1), "%\n")
```

### Cross-Party Heatmap

Which party pairs co-sponsor the most?

```{r cross-heatmap, exercise=TRUE, exercise.eval=FALSE, fig.height=5}
library(igraph)
proposers <- get_proposers()

props_22 <- proposers %>% filter(propose_date >= as.Date("2024-05-30"))
leads <- props_22 %>% filter(is_lead) %>% select(bill_id, lead = member_id)
cosponsors <- props_22 %>% filter(!is_lead) %>% select(bill_id, cosponsor = member_id)
edge_weights <- inner_join(leads, cosponsors, by = "bill_id") %>%
  filter(!is.na(lead), !is.na(cosponsor), lead != cosponsor) %>%
  count(lead, cosponsor, name = "weight")

g <- graph_from_data_frame(edge_weights, directed = FALSE)
leg_22 <- legislators %>% filter(assembly == 22) %>% distinct(member_id, .keep_all = TRUE)
V(g)$party <- leg_22$party[match(V(g)$name, leg_22$member_id)]

edge_df <- as_data_frame(g, what = "edges") %>%
  mutate(
    party_from = V(g)$party[match(from, V(g)$name)],
    party_to   = V(g)$party[match(to, V(g)$name)]
  ) %>%
  filter(!is.na(party_from), !is.na(party_to))

party_pair <- edge_df %>%
  filter(party_from %in% c("\uad6d\ubbfc\uc758\ud798", "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9", "\uc870\uad6d\ud601\uc2e0\ub2f9"),
         party_to %in% c("\uad6d\ubbfc\uc758\ud798", "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9", "\uc870\uad6d\ud601\uc2e0\ub2f9")) %>%
  count(party_from, party_to) %>%
  mutate(n_log = log10(n + 1))

ggplot(party_pair, aes(x = party_from, y = party_to, fill = n_log)) +
  geom_tile(color = "white") +
  geom_text(aes(label = scales::comma(n)), size = 4) +
  scale_fill_gradient(low = "lightyellow", high = "tomato", guide = "none") +
  labs(x = NULL, y = NULL,
       title = "Cross-Party Co-sponsorship Counts") +
  theme_minimal(base_size = 13)
```

If the diagonal (within-party) is much larger than off-diagonal (cross-party), co-sponsorship mainly occurs within party boundaries.

## 9. Network Visualization

```{r network-viz, exercise=TRUE, exercise.eval=FALSE, fig.height=6}
library(igraph)
proposers <- get_proposers()

props_22 <- proposers %>% filter(propose_date >= as.Date("2024-05-30"))
leads <- props_22 %>% filter(is_lead) %>% select(bill_id, lead = member_id)
cosponsors <- props_22 %>% filter(!is_lead) %>% select(bill_id, cosponsor = member_id)
edge_weights <- inner_join(leads, cosponsors, by = "bill_id") %>%
  filter(!is.na(lead), !is.na(cosponsor), lead != cosponsor) %>%
  count(lead, cosponsor, name = "weight")

g <- graph_from_data_frame(edge_weights, directed = FALSE)
leg_22 <- legislators %>% filter(assembly == 22) %>% distinct(member_id, .keep_all = TRUE)
V(g)$party <- leg_22$party[match(V(g)$name, leg_22$member_id)]

# Visualize two main parties only (full network is too complex)
main_parties <- c("\uad6d\ubbfc\uc758\ud798", "\ub354\ubd88\uc5b4\ubbfc\uc8fc\ub2f9")
keep <- V(g)$party %in% main_parties & !is.na(V(g)$party)
g_sub <- induced_subgraph(g, which(keep))

party_colors <- ifelse(V(g_sub)$party == "\uad6d\ubbfc\uc758\ud798", "#E61E2B", "#004EA2")

set.seed(42)
layout <- layout_with_fr(g_sub)

# Remove weak edges for visual clarity
g_viz <- delete_edges(g_sub, which(E(g_sub)$weight < 10))

plot(g_viz,
     vertex.color = party_colors,
     vertex.size = sqrt(degree(g_viz)) * 0.8,
     vertex.label = NA,
     edge.width = E(g_viz)$weight / 50,
     edge.color = rgb(0, 0, 0, 0.08),
     layout = layout,
     main = "22nd Assembly Co-sponsorship Network (PPP vs DPK)")

legend("bottomleft",
       legend = main_parties,
       col = c("#E61E2B", "#004EA2"),
       pch = 19, pt.cex = 1.5, bty = "n")
```

Are the two parties separated or mixed in the network? This visualization provides an intuitive grasp of cross-party co-sponsorship structure.

## Summary

| Concept | Meaning | Function |
|---------|---------|----------|
| Node / Edge | Unit of analysis / Relationship | `graph_from_data_frame()` |
| Degree centrality | Nodes with many connections | `degree()` |
| Betweenness centrality | Bridge nodes between groups | `betweenness()` |
| Community detection | Tightly connected subgroups | `cluster_louvain()` |
| Modularity | Clarity of community separation | `modularity()` |
| Cross-party ratio | Inter-party co-sponsorship | Edge attribute comparison |

## Exercises

### Exercise 1: Gender and Network Position

Compare mean betweenness centrality between female and male legislators. Is there a gender difference in network position?

```{r ex1, exercise=TRUE, exercise.eval=FALSE}

```

```{r ex1-hint}
# Build the network and centrality_df first, then
# add gender from leg_22 and group_by(gender)
```

```{r ex1-solution}
library(igraph)
proposers <- get_proposers()

props_22 <- proposers %>% filter(propose_date >= as.Date("2024-05-30"))
leads <- props_22 %>% filter(is_lead) %>% select(bill_id, lead = member_id)
cosponsors <- props_22 %>% filter(!is_lead) %>% select(bill_id, cosponsor = member_id)
edge_weights <- inner_join(leads, cosponsors, by = "bill_id") %>%
  filter(!is.na(lead), !is.na(cosponsor), lead != cosponsor) %>%
  count(lead, cosponsor, name = "weight")

g <- graph_from_data_frame(edge_weights, directed = FALSE)
leg_22 <- legislators %>% filter(assembly == 22) %>% distinct(member_id, .keep_all = TRUE)
V(g)$gender <- leg_22$gender[match(V(g)$name, leg_22$member_id)]
V(g)$label <- leg_22$name[match(V(g)$name, leg_22$member_id)]

data.frame(
  name = V(g)$label,
  gender = V(g)$gender,
  betweenness = betweenness(g, normalized = TRUE)
) %>%
  filter(!is.na(gender)) %>%
  group_by(gender) %>%
  summarise(
    n = n(),
    avg_betweenness = round(mean(betweenness), 6),
    median_betweenness = round(median(betweenness), 6)
  )
```

### Exercise 2: Alternative Community Detection

Use `cluster_walktrap()` instead of Louvain and compare the results. How do modularity and the number of communities differ?

```{r ex2, exercise=TRUE, exercise.eval=FALSE}

```

```{r ex2-hint}
# Build the graph g, then run both algorithms:
# comm_louvain <- cluster_louvain(g)
# comm_walktrap <- cluster_walktrap(g)
# Compare max(membership(...)) and modularity(...)
```

```{r ex2-solution}
library(igraph)
proposers <- get_proposers()

props_22 <- proposers %>% filter(propose_date >= as.Date("2024-05-30"))
leads <- props_22 %>% filter(is_lead) %>% select(bill_id, lead = member_id)
cosponsors <- props_22 %>% filter(!is_lead) %>% select(bill_id, cosponsor = member_id)
edge_weights <- inner_join(leads, cosponsors, by = "bill_id") %>%
  filter(!is.na(lead), !is.na(cosponsor), lead != cosponsor) %>%
  count(lead, cosponsor, name = "weight")

g <- graph_from_data_frame(edge_weights, directed = FALSE)

comm_louvain <- cluster_louvain(g)
comm_walktrap <- cluster_walktrap(g)

data.frame(
  Algorithm = c("Louvain", "Walktrap"),
  Communities = c(max(membership(comm_louvain)), max(membership(comm_walktrap))),
  Modularity = c(round(modularity(comm_louvain), 3), round(modularity(comm_walktrap), 3))
)
```

### Exercise 3: 21st vs 22nd Assembly Comparison

Build the 21st Assembly co-sponsorship network and compare with the 22nd. Report at least one of: density, average degree, or cross-party ratio.

```{r ex3, exercise=TRUE, exercise.eval=FALSE}

```

```{r ex3-hint}
# Filter proposers by propose_date for 21st Assembly:
# propose_date >= "2020-05-30" & propose_date < "2024-05-30"
# Build network same way as 22nd, then compare metrics
```

```{r ex3-solution}
library(igraph)
proposers <- get_proposers()

# 21st Assembly
props_21 <- proposers %>%
  filter(propose_date >= as.Date("2020-05-30"),
         propose_date < as.Date("2024-05-30"))

leads_21 <- props_21 %>% filter(is_lead) %>% select(bill_id, lead = member_id)
cos_21 <- props_21 %>% filter(!is_lead) %>% select(bill_id, cosponsor = member_id)

ew_21 <- inner_join(leads_21, cos_21, by = "bill_id") %>%
  filter(!is.na(lead), !is.na(cosponsor), lead != cosponsor) %>%
  count(lead, cosponsor, name = "weight")

g_21 <- graph_from_data_frame(ew_21, directed = FALSE)

# 22nd Assembly
props_22 <- proposers %>% filter(propose_date >= as.Date("2024-05-30"))
leads_22 <- props_22 %>% filter(is_lead) %>% select(bill_id, lead = member_id)
cos_22 <- props_22 %>% filter(!is_lead) %>% select(bill_id, cosponsor = member_id)

ew_22 <- inner_join(leads_22, cos_22, by = "bill_id") %>%
  filter(!is.na(lead), !is.na(cosponsor), lead != cosponsor) %>%
  count(lead, cosponsor, name = "weight")

g_22 <- graph_from_data_frame(ew_22, directed = FALSE)

data.frame(
  Assembly = c("21st", "22nd"),
  Nodes = c(vcount(g_21), vcount(g_22)),
  Edges = c(ecount(g_21), ecount(g_22)),
  Density = c(round(graph.density(g_21), 4), round(graph.density(g_22), 4)),
  Avg_Degree = c(round(mean(degree(g_21)), 1), round(mean(degree(g_22)), 1))
)
```
